---
title: "Data Preparation and Validation"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# Executive Summary

## Background

This notebook loads and validates the pre-processed drug response data (`ccle.Rda` and
`Sanger.Rdata`) that are used throughout this project. These files were generated from
raw CCLE/GDSC data using code archived in `archive/code_legacy/30-replication.Rmd`.

## Data Sources

- **ccle.Rda**: Contains predictor matrix `x` (genomic features), response matrix `y`
  (drug sensitivity), and cell line selector `cls`
- **Sanger.Rdata**: Contains the full Sanger/GDSC drug response data with tissue annotations

See `doc/DATA_PROVENANCE.md` for detailed documentation of how these files were created.

# Setup

```{r load-packages}
library(here)
library(tidyverse)
```

```{r load-paths}
source(here("methods", "00_paths.R"))
```

# Load Data

## Load ccle.Rda

```{r load-ccle}
ccle_file <- file.path(paths$clean, "ccle.Rda")

if (!file.exists(ccle_file)) {
  stop("ccle.Rda not found in ", paths$clean,
       "\nSee doc/DATA_PROVENANCE.md for instructions")
}

load(ccle_file)
message("Loaded ccle.Rda successfully")
```

## Load Sanger.Rdata

```{r load-sanger}
sanger_file <- file.path(paths$clean, "Sanger.Rdata")

if (!file.exists(sanger_file)) {
  stop("Sanger.Rdata not found in ", paths$clean,
       "\nSee doc/DATA_PROVENANCE.md for instructions")
}

load(sanger_file)
message("Loaded Sanger.Rdata successfully")
```

# Data Validation

## Predictor Matrix (x)

```{r validate-x}
cat("Predictor matrix dimensions:\n")
cat(sprintf("  Samples (cell lines): %d\n", nrow(x)))
cat(sprintf("  Features: %d\n", ncol(x)))

# Feature type breakdown
feature_types <- data.frame(
  Type = c("Expression", "Mutation", "Copy Number", "Protein (RPPA)", "Other"),
  Count = c(
    sum(grepl("_Expr$", colnames(x))),
    sum(grepl("_mut\\.", colnames(x))),
    sum(grepl("_CN$", colnames(x))),
    sum(!grepl("_Expr$|_mut\\.|_CN$", colnames(x)) &
        grepl("_|^[A-Z]+$", colnames(x))),
    0  # Will calculate as remainder
  )
)
feature_types$Count[5] <- ncol(x) - sum(feature_types$Count[1:4])

knitr::kable(feature_types, caption = "Feature Type Breakdown")
```

## Response Matrix (y)

```{r validate-y}
cat("Response matrix dimensions:\n")
cat(sprintf("  Samples (cell lines): %d\n", nrow(y)))
cat(sprintf("  Drugs: %d\n", ncol(y)))

# Missing data summary
missing_pct <- mean(is.na(y)) * 100
cat(sprintf("\nMissing values: %.1f%%\n", missing_pct))

# Samples per drug
samples_per_drug <- colSums(!is.na(y))
cat(sprintf("\nSamples per drug:\n"))
cat(sprintf("  Min: %d\n", min(samples_per_drug)))
cat(sprintf("  Median: %d\n", median(samples_per_drug)))
cat(sprintf("  Max: %d\n", max(samples_per_drug)))
```

## Sanger Object

```{r validate-sanger}
cat("Sanger object contents:\n")
cat(sprintf("  Drug response (dat): %d x %d\n", nrow(Sanger$dat), ncol(Sanger$dat)))
cat(sprintf("  IC50 values: %d x %d\n", nrow(Sanger$ic50), ncol(Sanger$ic50)))
cat(sprintf("  Cell lines with data: %d\n", sum(Sanger$cls)))

# Tissue distribution
if (!is.null(Sanger$tissue)) {
  tissue_counts <- sort(table(Sanger$tissue), decreasing = TRUE)
  cat(sprintf("\nTissue types: %d\n", length(tissue_counts)))
  cat("\nTop 10 tissues:\n")
  print(head(tissue_counts, 10))
} else {
  cat("\nNote: Tissue information not available in Sanger object\n")
}
```

# Cell Line Overlap

```{r cell-line-overlap}
# Check overlap between x and Sanger
x_samples <- rownames(x)
sanger_samples <- rownames(Sanger$dat)

common_samples <- intersect(x_samples, sanger_samples)
cat(sprintf("Cell lines in predictor matrix: %d\n", length(x_samples)))
cat(sprintf("Cell lines in Sanger: %d\n", length(sanger_samples)))
cat(sprintf("Common cell lines: %d\n", length(common_samples)))
```

# Drug List

```{r drug-list}
drugs <- colnames(Sanger$dat)
cat(sprintf("Total drugs: %d\n", length(drugs)))
cat("\nFirst 20 drugs:\n")
print(head(drugs, 20))
```

# Data Quality Checks

```{r quality-checks}
# Check for any all-NA rows or columns in predictors
na_rows <- rowSums(is.na(x)) == ncol(x)
na_cols <- colSums(is.na(x)) == nrow(x)

cat("Quality checks:\n")
cat(sprintf("  All-NA rows in x: %d\n", sum(na_rows)))
cat(sprintf("  All-NA columns in x: %d\n", sum(na_cols)))

# Check for zero-variance features
zero_var <- apply(x, 2, var, na.rm = TRUE) == 0
cat(sprintf("  Zero-variance features: %d\n", sum(zero_var, na.rm = TRUE)))
```

# Summary

The data has been loaded and validated successfully. Key statistics:

- **`r nrow(x)`** cell lines with **`r ncol(x)`** genomic features
- **`r ncol(Sanger$dat)`** drugs tested
- **`r length(common_samples)`** cell lines with both genomic and drug response data
- **`r if(!is.null(Sanger$tissue)) length(unique(Sanger$tissue)) else "N/A"`** tissue types

The data is ready for downstream analysis.

# Session Info

```{r session-info}
sessionInfo()
```
