---
title: "Cross-Validation: Wrong vs Right Feature Selection"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(

  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# Executive Summary

## Background

This is the **main methodological study** of the project. We compare two approaches
to cross-validation for elastic net drug response prediction:

1. **Wrong (Published)**: Feature selection on FULL data BEFORE CV (data leakage)
2. **Right (Correct)**: Feature selection WITHIN each CV fold (no leakage)

The "wrong" approach is what was used in published papers (Barrentina 2012, Ghandi 2019).
It leads to overly optimistic performance estimates due to information leaking from
test to train.

## Key Methodological Fix

**Standard Error vs Standard Deviation Bug**: The original implementations reported
SD of CV errors instead of SE. The correct formula is:

$$SE = \frac{SD}{\sqrt{K}}$$

where $K$ is the number of folds. This implementation uses the correct SE.

# Setup

```{r load-packages}
library(here)
library(tidyverse)
library(glmnet)
library(caret)
```

```{r load-methods}
source(here("methods", "00_paths.R"))
source(here("methods", "01_data_utils.R"))
source(here("methods", "02_cv_utils.R"))
source(here("methods", "04_plot_utils.R"))
```

# Configuration

```{r config}
# CV parameters
CV_NFOLDS <- 5
BOOTSTRAP_NREPS <- 200
CORRELATION_THRESHOLD <- 0.1
VARIANCE_THRESHOLD <- 0.01
ALPHA_GRID <- seq(0.2, 1.0, by = 0.1)
GLOBAL_SEED <- 42

# Analysis parameters
MIN_SAMPLES <- 50  # Minimum samples per drug
N_DRUGS <- NULL    # NULL = all drugs, or set to integer for testing

# Output
results_dir <- file.path(paths$results, "cv_results")
if (!dir.exists(results_dir)) dir.create(results_dir, recursive = TRUE)
```

# Load Data

```{r load-data}
data <- load_main_data("Sanger")
print_data_summary(data)
```

# Select Drugs for Analysis

```{r select-drugs}
drugs <- get_drug_list(data)

# Filter to drugs with enough samples
valid_drugs <- sapply(drugs, function(d) {
  drug_data <- prepare_drug_data(data, d, min_samples = MIN_SAMPLES)
  !is.null(drug_data)
})

drugs_to_analyze <- drugs[valid_drugs]

if (!is.null(N_DRUGS)) {
  drugs_to_analyze <- drugs_to_analyze[1:min(N_DRUGS, length(drugs_to_analyze))]
}

cat(sprintf("Analyzing %d drugs (of %d total)\n",
            length(drugs_to_analyze), length(drugs)))
```

# Run Comparison

```{r load-precomputed-results}
results_file <- file.path(paths$scratch, "cv_wrong_vs_right_results.Rda")

if (!file.exists(results_file)) {
  stop(
    "Pre-computed results not found.\n",
    "Run: Rscript methods/run_cv_wrong_vs_right.R\n",
    "Or:  TEST_MODE=TRUE Rscript methods/run_cv_wrong_vs_right.R"
  )
}

load(results_file)

if (!exists("cv_results")) {
  stop("Expected object 'cv_results' in results file.")
}

cat("Loaded pre-computed results\n")
cat("  Drugs:", nrow(cv_results), "\n")
cat("  Successful:", sum(cv_results$status == "success"), "\n\n")

success_cv <- cv_results %>% filter(status == "success")
```

# Results Summary

```{r summary-table}
# Create summary table
summary_df <- success_cv %>%
  transmute(
    Drug = drug,
    N_samples = n_samples,
    MSE_wrong = mse_wrong,
    MSE_right = mse_right,
    MSE_increase_pct = inflation_pct,
    Alpha_wrong = alpha_wrong,
    Alpha_right = alpha_right,
    Jaccard = jaccard
  )

# Display summary
knitr::kable(
  summary_df %>%
    arrange(desc(MSE_increase_pct)) %>%
    head(20),
  digits = 3,
  caption = "Top 20 drugs by MSE increase (wrong -> right)"
)
```

```{r summary-stats}
cat("Overall Statistics:\n")
cat(sprintf("  Drugs analyzed: %d\n", nrow(summary_df)))
cat(sprintf("  Mean MSE wrong: %.4f (SD: %.4f)\n",
            mean(summary_df$MSE_wrong), sd(summary_df$MSE_wrong)))
cat(sprintf("  Mean MSE right: %.4f (SD: %.4f)\n",
            mean(summary_df$MSE_right), sd(summary_df$MSE_right)))
cat(sprintf("  Mean MSE increase: %.1f%%\n",
            mean(summary_df$MSE_increase_pct)))
cat(sprintf("  Mean Jaccard similarity: %.3f\n",
            mean(summary_df$Jaccard, na.rm = TRUE)))
```

# Visualizations

## MSE Comparison

```{r plot-mse-comparison, fig.height=6}
ggplot(summary_df, aes(x = MSE_wrong, y = MSE_right)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(alpha = 0.6, size = 2) +
  labs(
    title = "Cross-Validation MSE: Wrong vs Right",
    subtitle = "Points above diagonal = correct CV gives higher (more realistic) error",
    x = "MSE (Wrong CV - with leakage)",
    y = "MSE (Right CV - no leakage)"
  ) +
  theme_minimal() +
  theme(text = element_text(size = 12))
```

## Jaccard Similarity Distribution

```{r plot-jaccard, fig.height=5}
ggplot(summary_df, aes(x = Jaccard)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white", alpha = 0.8) +
  geom_vline(xintercept = mean(summary_df$Jaccard, na.rm = TRUE),
             linetype = "dashed", color = "red", size = 1) +
  labs(
    title = "Feature Selection Stability: Jaccard Similarity",
    subtitle = "Overlap between features selected by wrong vs right CV",
    x = "Jaccard Index",
    y = "Count"
  ) +
  theme_minimal()
```

## MSE Increase Distribution

```{r plot-mse-increase, fig.height=5}
ggplot(summary_df, aes(x = MSE_increase_pct)) +
  geom_histogram(bins = 30, fill = "coral", color = "white", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "MSE Increase When Using Correct CV",
    subtitle = "Positive values = correct CV reveals worse (more realistic) performance",
    x = "MSE Increase (%)",
    y = "Count"
  ) +
  theme_minimal()
```

# Save Summary Table

```{r save-summary}
summary_file <- file.path(results_dir, "cv_comparison_summary.csv")
write_csv(summary_df, summary_file)
cat(sprintf("Summary saved to: %s\n", summary_file))
```

# Conclusions

1. **Data leakage matters**: The "wrong" CV approach systematically underestimates
   prediction error by an average of `r sprintf("%.1f%%", mean(summary_df$MSE_increase_pct))`.

2. **Feature instability**: The low Jaccard similarity (`r sprintf("%.2f", mean(summary_df$Jaccard, na.rm = TRUE))` mean)
   indicates that the features selected depend heavily on whether proper CV is used.

3. **Standard Error fix**: This analysis uses the correct SE formula (SD/sqrt(K))
   rather than SD, which was the bug in prior implementations.

# Session Info

```{r session-info}
sessionInfo()
```
