---
title: "Aggregate CV Results Analysis"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# Executive Summary

This notebook aggregates and analyzes the results from the wrong vs right CV comparison
(experiment 02). It provides summary statistics, visualizations, and statistical tests
to quantify the impact of data leakage in feature selection.

# Setup

```{r load-packages}
library(here)
library(tidyverse)
library(knitr)
library(ggpubr)
```

```{r load-paths}
source(here("methods", "00_paths.R"))
source(here("methods", "04_plot_utils.R"))
```

# Load Results

```{r load-results}
results_file <- file.path(paths$results, "cv_results", "cv_comparison_full.Rda")

if (!file.exists(results_file)) {
  stop("Results file not found: ", results_file,
       "\nRun 02_cv_wrong_vs_right.Rmd first")
}

load(results_file)  # Loads all_results
cat(sprintf("Loaded results for %d drugs\n", length(all_results)))
```

# Summary Statistics

```{r create-summary}
# Create comprehensive summary
summary_df <- map_dfr(all_results, function(res) {
  tibble(
    Drug = res$drug,
    N_samples = res$n_samples,
    N_features = res$n_features,

    # Wrong CV
    MSE_wrong = res$comparison$mse_wrong,
    Alpha_wrong = res$comparison$alpha_wrong,
    Lambda_wrong = res$comparison$lambda_wrong,
    N_features_wrong = res$comparison$n_features_wrong,

    # Right CV
    MSE_right = res$comparison$mse_right,
    Alpha_right = res$comparison$alpha_right,
    Lambda_right = res$comparison$lambda_right,
    N_features_right = res$comparison$n_features_right,

    # Comparison
    MSE_diff = res$comparison$mse_diff,
    MSE_increase_pct = res$comparison$mse_pct_increase,
    Jaccard = res$comparison$jaccard
  )
})
```

## Overall Statistics

```{r overall-stats}
stats_table <- tribble(
  ~Metric, ~Value,
  "Total drugs analyzed", nrow(summary_df),
  "Mean samples per drug", sprintf("%.1f", mean(summary_df$N_samples)),
  "Mean MSE (wrong CV)", sprintf("%.4f", mean(summary_df$MSE_wrong)),
  "Mean MSE (right CV)", sprintf("%.4f", mean(summary_df$MSE_right)),
  "Mean MSE increase (%)", sprintf("%.1f%%", mean(summary_df$MSE_increase_pct)),
  "Median MSE increase (%)", sprintf("%.1f%%", median(summary_df$MSE_increase_pct)),
  "Mean Jaccard similarity", sprintf("%.3f", mean(summary_df$Jaccard, na.rm = TRUE)),
  "Drugs where right > wrong", sprintf("%d (%.1f%%)",
    sum(summary_df$MSE_right > summary_df$MSE_wrong),
    100 * mean(summary_df$MSE_right > summary_df$MSE_wrong))
)

kable(stats_table, caption = "Summary Statistics")
```

## Statistical Test

```{r paired-test}
# Paired t-test for MSE difference
t_result <- t.test(summary_df$MSE_right, summary_df$MSE_wrong, paired = TRUE)

cat("Paired t-test: MSE_right vs MSE_wrong\n")
cat(sprintf("  t-statistic: %.3f\n", t_result$statistic))
cat(sprintf("  p-value: %.2e\n", t_result$p.value))
cat(sprintf("  95%% CI: [%.4f, %.4f]\n",
            t_result$conf.int[1], t_result$conf.int[2]))
```

# Visualizations

## MSE Comparison Scatter

```{r scatter-mse, fig.height=7}
p1 <- ggplot(summary_df, aes(x = MSE_wrong, y = MSE_right)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50", size = 1) +
  geom_point(aes(color = MSE_increase_pct), alpha = 0.7, size = 3) +
  scale_color_gradient2(low = "blue", mid = "gray", high = "red", midpoint = 0,
                        name = "MSE Increase (%)") +
  labs(
    title = "Cross-Validation MSE: Wrong vs Right",
    subtitle = sprintf("n = %d drugs | Paired t-test p = %.2e", nrow(summary_df), t_result$p.value),
    x = "MSE (Wrong CV - with data leakage)",
    y = "MSE (Right CV - no leakage)"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")

print(p1)
```

## Distribution of MSE Increase

```{r hist-increase, fig.height=5}
p2 <- ggplot(summary_df, aes(x = MSE_increase_pct)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "steelblue",
                 color = "white", alpha = 0.8) +
  geom_density(color = "darkblue", size = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = mean(summary_df$MSE_increase_pct),
             linetype = "solid", color = "darkgreen", size = 1) +
  annotate("text", x = mean(summary_df$MSE_increase_pct), y = Inf,
           label = sprintf("Mean: %.1f%%", mean(summary_df$MSE_increase_pct)),
           vjust = 2, hjust = -0.1, color = "darkgreen") +
  labs(
    title = "Distribution of MSE Increase (Right vs Wrong)",
    subtitle = "Positive = correct CV reveals worse performance (expected)",
    x = "MSE Increase (%)",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)

print(p2)
```

## Jaccard Similarity Distribution

```{r hist-jaccard, fig.height=5}
p3 <- ggplot(summary_df, aes(x = Jaccard)) +
  geom_histogram(aes(y = ..density..), bins = 25, fill = "coral",
                 color = "white", alpha = 0.8) +
  geom_density(color = "darkred", size = 1) +
  geom_vline(xintercept = mean(summary_df$Jaccard, na.rm = TRUE),
             linetype = "dashed", color = "darkgreen", size = 1) +
  labs(
    title = "Feature Selection Stability (Jaccard Index)",
    subtitle = sprintf("Mean: %.3f | 1.0 = identical features, 0 = no overlap",
                       mean(summary_df$Jaccard, na.rm = TRUE)),
    x = "Jaccard Similarity",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)

print(p3)
```

## MSE vs Sample Size

```{r mse-vs-n, fig.height=5}
p4 <- ggplot(summary_df, aes(x = N_samples, y = MSE_increase_pct)) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "loess", se = TRUE, color = "blue") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "MSE Increase vs Sample Size",
    subtitle = "Does the effect of data leakage depend on sample size?",
    x = "Number of Samples",
    y = "MSE Increase (%)"
  ) +
  theme_minimal(base_size = 14)

print(p4)
```

# Top/Bottom Drugs

## Most Affected by Leakage

```{r top-affected}
kable(
  summary_df %>%
    arrange(desc(MSE_increase_pct)) %>%
    select(Drug, N_samples, MSE_wrong, MSE_right, MSE_increase_pct, Jaccard) %>%
    head(15),
  digits = 3,
  caption = "Top 15 Drugs Most Affected by Data Leakage"
)
```

## Least Affected by Leakage

```{r bottom-affected}
kable(
  summary_df %>%
    arrange(MSE_increase_pct) %>%
    select(Drug, N_samples, MSE_wrong, MSE_right, MSE_increase_pct, Jaccard) %>%
    head(15),
  digits = 3,
  caption = "Top 15 Drugs Least Affected by Data Leakage"
)
```

# Save Outputs

```{r save-outputs}
# Save summary table
write_csv(summary_df, file.path(paths$results, "tables", "cv_aggregate_summary.csv"))

# Save key figures
ggsave(file.path(paths$results, "figures", "cv_mse_comparison.pdf"),
       p1, width = 10, height = 7)
ggsave(file.path(paths$results, "figures", "cv_mse_increase_dist.pdf"),
       p2, width = 8, height = 5)
ggsave(file.path(paths$results, "figures", "cv_jaccard_dist.pdf"),
       p3, width = 8, height = 5)

cat("Outputs saved to results/tables/ and results/figures/\n")
```

# Conclusions

1. **Data leakage inflates performance**: Across `r nrow(summary_df)` drugs, correct CV
   shows `r sprintf("%.1f%%", mean(summary_df$MSE_increase_pct))` higher MSE on average.

2. **Feature selection is unstable**: Mean Jaccard similarity of
   `r sprintf("%.3f", mean(summary_df$Jaccard, na.rm = TRUE))` indicates the features
   selected depend heavily on whether data leakage is present.

3. **Effect is robust**: The paired t-test shows the difference is highly significant
   (p = `r sprintf("%.2e", t_result$p.value)`).

# Session Info

```{r session-info}
sessionInfo()
```
