---
title: "Feature Selection Stability Analysis"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 150
)
```

# Executive Summary

This notebook analyzes the stability of feature selection across different modeling
approaches:

1. **Wrong vs Right CV**: How much do selected features change when proper cross-validation
   is applied (per-fold feature selection)?

2. **EN vs DSEN**: How does incorporating tissue structure affect which features are
   selected and how stably?

We use Jaccard similarity and bootstrap selection frequencies to quantify feature
stability across methods.

## Key Findings Preview

- **Wrong vs Right CV**: Feature overlap (Jaccard) is typically low, showing that
  data leakage changes which features appear predictive
- **EN vs DSEN**: DSEN selects 42% fewer features on average, with higher selection
  stability due to the structured penalty

# Setup

```{r load-packages}
library(here)
library(tidyverse)
library(pheatmap)
library(ggrepel)
library(patchwork)
library(scales)
library(knitr)
```

```{r load-paths}
source(here("methods", "00_paths.R"))
source(here("methods", "04_plot_utils.R"))
```

```{r output-dirs}
fig_dir <- file.path(paths$results, "figures")
table_dir <- file.path(paths$results, "tables")
for (d in c(fig_dir, table_dir)) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Load Results

## Wrong vs Right CV Results

```{r load-cv-results}
# Try multiple locations for CV results
cv_results_files <- c(
  file.path(paths$scratch, "feature_stability_results.Rda"),
  here("results", "cv_results", "feature_stability_results.Rda")
)

cv_file <- cv_results_files[file.exists(cv_results_files)][1]

if (!is.na(cv_file)) {
  load(cv_file)

  # Filter to successful results
  cv_stability_results <- Filter(function(res) {
    !is.null(res) && (is.null(res$status) || res$status == "success")
  }, all_results)

  rm(all_results)
  has_cv_stability <- TRUE
  cat(sprintf("Loaded CV stability results for %d drugs\n", length(cv_stability_results)))
} else {
  has_cv_stability <- FALSE
  cat("CV stability results not found. Wrong vs Right analysis will be skipped.\n")
  cat("Run: Rscript methods/run_feature_stability.R\n")
}
```

## DSEN Bootstrap Results

```{r load-dsen-results}
dsen_bootstrap_file <- here("results", "dsen_results", "dsen_bootstrap_results.Rda")

if (file.exists(dsen_bootstrap_file)) {
  load(dsen_bootstrap_file)
  dsen_bootstrap_results <- all_results
  rm(all_results)
  has_dsen_bootstrap <- TRUE
  cat(sprintf("Loaded DSEN bootstrap results for %d drugs\n", length(dsen_bootstrap_results)))
} else {
  has_dsen_bootstrap <- FALSE
  cat("DSEN bootstrap results not found. EN vs DSEN analysis will be limited.\n")
  cat("Run: Rscript methods/run_dsen_bootstrap.R\n")
}

# Also load DSEN summary for context
dsen_summary_file <- here("results", "dsen_results", "dsen_summary.csv")
if (file.exists(dsen_summary_file)) {
  dsen_summary <- read_csv(dsen_summary_file, show_col_types = FALSE)
  cat("Loaded DSEN summary CSV\n")
}
```

# Part 1: Wrong vs Right CV Feature Stability

## Rationale

When feature selection is performed incorrectly (on full data before CV split), the
selected features are "optimized" to the test set, creating artificial consistency.
Correct CV (per-fold feature selection) reveals the true instability of feature
selection. Low Jaccard similarity between wrong and right CV selected features
indicates that data leakage changes biomarker identification.

```{r cv-stability-check}
if (!has_cv_stability) {
  cat("Skipping Wrong vs Right CV analysis (results not available)\n")
}
```

## Jaccard Similarity Distribution

```{r jaccard-overall, eval=has_cv_stability}
jaccard_df <- map_dfr(cv_stability_results, function(res) {
  tibble(
    Drug = res$drug,
    Jaccard = res$comparison$jaccard,
    N_samples = res$n_samples,
    N_wrong = res$comparison$n_features_wrong,
    N_right = res$comparison$n_features_right
  )
})

# Summary statistics
cat("=== Jaccard Similarity (Wrong vs Right CV) ===\n")
cat(sprintf("  Mean: %.3f\n", mean(jaccard_df$Jaccard, na.rm = TRUE)))
cat(sprintf("  Median: %.3f\n", median(jaccard_df$Jaccard, na.rm = TRUE)))
cat(sprintf("  SD: %.3f\n", sd(jaccard_df$Jaccard, na.rm = TRUE)))
cat(sprintf("  Range: [%.3f, %.3f]\n",
            min(jaccard_df$Jaccard, na.rm = TRUE),
            max(jaccard_df$Jaccard, na.rm = TRUE)))
cat(sprintf("  Drugs with Jaccard = 0: %d (%.1f%%)\n",
            sum(jaccard_df$Jaccard == 0, na.rm = TRUE),
            100 * mean(jaccard_df$Jaccard == 0, na.rm = TRUE)))
```

```{r jaccard-histogram, eval=has_cv_stability, fig.height=5}
mean_jaccard <- mean(jaccard_df$Jaccard, na.rm = TRUE)
zero_pct <- 100 * mean(jaccard_df$Jaccard == 0, na.rm = TRUE)

p_jaccard <- ggplot(jaccard_df, aes(x = Jaccard)) +
  geom_histogram(bins = 25, fill = "#377EB8", color = "white", alpha = 0.8) +
  geom_vline(xintercept = mean_jaccard,
             linetype = "dashed", color = "#E41A1C", linewidth = 1) +
  annotate("text", x = mean_jaccard + 0.03, y = Inf,
           label = sprintf("Mean = %.3f", mean_jaccard),
           vjust = 2, hjust = 0, color = "#E41A1C", size = 4) +
  annotate("text", x = 0.05, y = Inf,
           label = sprintf("%.1f%% with\nJaccard = 0", zero_pct),
           vjust = 3, hjust = 0, color = "#E41A1C", size = 3.5) +
  labs(
    title = "Feature Overlap Between Wrong and Right CV",
    subtitle = "Jaccard = |intersection| / |union| of selected features",
    x = "Jaccard Similarity",
    y = "Number of Drugs"
  ) +
  scale_x_continuous(limits = c(0, 1)) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank())

print(p_jaccard)

ggsave(file.path(fig_dir, "feature_jaccard_wrongvright.pdf"),
       p_jaccard, width = 8, height = 5)
```

## Interpretation

The low mean Jaccard similarity (`r if(has_cv_stability) sprintf("%.3f", mean_jaccard) else "N/A"`)
demonstrates that wrong and right CV select substantially different feature sets.
This has serious implications for biomarker discovery: features identified as
"predictive" using incorrect methodology may not replicate when proper CV is applied.

## Drugs with Lowest Feature Overlap

```{r low-jaccard, eval=has_cv_stability}
low_jaccard <- jaccard_df %>%
  arrange(Jaccard) %>%
  head(15)

kable(low_jaccard, digits = 3,
      caption = "Drugs with Lowest Feature Overlap (Most Different Selections)")
```

## Feature Count Comparison: Wrong vs Right

```{r feature-count-cv, eval=has_cv_stability, fig.height=5}
p_count_cv <- ggplot(jaccard_df, aes(x = N_wrong, y = N_right)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(alpha = 0.6, size = 2, color = "#666666") +
  geom_smooth(method = "lm", se = TRUE, color = "#377EB8", fill = "#377EB8", alpha = 0.2) +
  labs(
    title = "Feature Count: Wrong vs Right CV",
    subtitle = "Wrong CV typically selects more features",
    x = "Features (Wrong CV)",
    y = "Features (Right CV)"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank())

print(p_count_cv)

cat(sprintf("\nMean features (Wrong CV): %.1f\n", mean(jaccard_df$N_wrong, na.rm = TRUE)))
cat(sprintf("Mean features (Right CV): %.1f\n", mean(jaccard_df$N_right, na.rm = TRUE)))
```

# Part 2: EN vs DSEN Feature Stability

## Rationale

DSEN (Data Shared Elastic Net) incorporates tissue structure through a penalty that
distinguishes between shared (pan-cancer) and tissue-specific features. We expect
DSEN to provide:

1. **Sparser models**: By leveraging tissue structure, fewer features need to be selected
2. **More stable selection**: The structured penalty should reduce spurious selections
3. **Interpretable decomposition**: Features split into shared vs tissue-specific

```{r dsen-check}
if (!has_dsen_bootstrap) {
  cat("Skipping EN vs DSEN analysis (bootstrap results not available)\n")
}
```

## Feature Count Comparison: EN vs DSEN

```{r feature-counts-dsen, eval=has_dsen_bootstrap}
# Extract feature counts from bootstrap results
dsen_feature_df <- map_dfr(dsen_bootstrap_results, function(r) {
  tibble(
    drug = r$drug,
    n_en = sum(r$en_bootstrap$cnt > 0),
    n_dsen = sum(r$dsen_bootstrap$cnt > 0),
    n_bootstrap = r$en_bootstrap$nboot
  )
}) %>%
  mutate(
    reduction = n_en - n_dsen,
    reduction_pct = 100 * reduction / n_en,
    dsen_more_sparse = n_dsen < n_en
  )

cat("=== Feature Count Summary (EN vs DSEN) ===\n")
cat(sprintf("Mean EN features: %.1f\n", mean(dsen_feature_df$n_en)))
cat(sprintf("Mean DSEN features: %.1f\n", mean(dsen_feature_df$n_dsen)))
cat(sprintf("Mean reduction: %.1f (%.1f%%)\n",
            mean(dsen_feature_df$reduction),
            mean(dsen_feature_df$reduction_pct)))
cat(sprintf("DSEN more sparse for %.1f%% of drugs\n",
            100 * mean(dsen_feature_df$dsen_more_sparse)))
```

```{r feature-scatter-dsen, eval=has_dsen_bootstrap, fig.height=6}
p_scatter_dsen <- ggplot(dsen_feature_df, aes(x = n_en, y = n_dsen)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = dsen_more_sparse), alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = TRUE, color = "#FF7F00", fill = "#FF7F00", alpha = 0.2) +
  scale_color_manual(
    values = c("TRUE" = "#FF7F00", "FALSE" = "#4DAF4A"),
    labels = c("TRUE" = "DSEN More Sparse", "FALSE" = "EN More Sparse"),
    name = NULL
  ) +
  labs(
    title = "Feature Selection: EN vs DSEN (Bootstrap)",
    subtitle = sprintf("DSEN selects %.0f%% fewer features on average | Bootstrap = %d iterations",
                      mean(dsen_feature_df$reduction_pct),
                      dsen_feature_df$n_bootstrap[1]),
    x = "Number of Features (Standard EN)",
    y = "Number of Features (DSEN)"
  ) +
  coord_fixed() +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = c(0.8, 0.2)
  )

print(p_scatter_dsen)

ggsave(file.path(fig_dir, "feature_count_en_vs_dsen.pdf"),
       p_scatter_dsen, width = 8, height = 6)
```

## Feature Reduction Distribution

```{r reduction-dist, eval=has_dsen_bootstrap, fig.height=5}
p_reduction <- ggplot(dsen_feature_df, aes(x = reduction_pct)) +
  geom_histogram(bins = 35, fill = "#FF7F00", color = "white", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray30") +
  geom_vline(xintercept = mean(dsen_feature_df$reduction_pct),
             linetype = "solid", color = "#984EA3", linewidth = 1) +
  annotate("text", x = mean(dsen_feature_df$reduction_pct) + 2, y = Inf,
           label = sprintf("Mean: %.1f%%", mean(dsen_feature_df$reduction_pct)),
           vjust = 2, hjust = 0, color = "#984EA3", size = 4) +
  labs(
    title = "Feature Reduction: DSEN vs Standard EN",
    subtitle = "Positive = DSEN selects fewer features",
    x = "Feature Reduction (%)",
    y = "Number of Drugs"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank())

print(p_reduction)

ggsave(file.path(fig_dir, "feature_reduction_dsen.pdf"),
       p_reduction, width = 8, height = 5)
```

## Interpretation

DSEN consistently selects fewer features than standard EN (42% fewer on average).
This sparsity advantage comes from the structured penalty that encourages either
shared effects (across all tissues) or tissue-specific effects, but discourages
noisy, inconsistent selections.

# Part 3: Feature Selection Stability Analysis

## Rationale

Beyond counting features, we want to understand selection stability across bootstrap
resamples. A feature selected in 90/100 bootstraps is more reliably associated with
drug response than one selected in 20/100 bootstraps.

## Bootstrap Selection Frequency Distribution

```{r stability-analysis, eval=has_dsen_bootstrap}
# Analyze selection frequency distributions
stability_analysis <- map_dfr(dsen_bootstrap_results, function(r) {
  # EN bootstrap frequencies
  en_freq <- r$en_bootstrap$cnt / r$en_bootstrap$nboot
  en_selected <- en_freq[en_freq > 0]

  # DSEN bootstrap frequencies
  dsen_freq <- r$dsen_bootstrap$cnt / r$dsen_bootstrap$nboot
  dsen_selected <- dsen_freq[dsen_freq > 0]

  tibble(
    drug = r$drug,
    # EN statistics
    en_n_selected = length(en_selected),
    en_mean_freq = mean(en_selected),
    en_median_freq = median(en_selected),
    en_n_stable = sum(en_freq >= 0.8),   # Selected in 80%+ bootstraps
    en_n_unstable = sum(en_freq > 0 & en_freq < 0.5),  # Selected in <50%
    # DSEN statistics
    dsen_n_selected = length(dsen_selected),
    dsen_mean_freq = mean(dsen_selected),
    dsen_median_freq = median(dsen_selected),
    dsen_n_stable = sum(dsen_freq >= 0.8),
    dsen_n_unstable = sum(dsen_freq > 0 & dsen_freq < 0.5)
  )
})

cat("\n=== Selection Stability Summary ===\n")
cat("\nStandard EN:\n")
cat(sprintf("  Mean features with freq >= 0.8: %.1f\n", mean(stability_analysis$en_n_stable)))
cat(sprintf("  Mean features with freq < 0.5: %.1f\n", mean(stability_analysis$en_n_unstable)))
cat(sprintf("  Mean selection frequency: %.3f\n", mean(stability_analysis$en_mean_freq)))

cat("\nDSEN:\n")
cat(sprintf("  Mean features with freq >= 0.8: %.1f\n", mean(stability_analysis$dsen_n_stable)))
cat(sprintf("  Mean features with freq < 0.5: %.1f\n", mean(stability_analysis$dsen_n_unstable)))
cat(sprintf("  Mean selection frequency: %.3f\n", mean(stability_analysis$dsen_mean_freq)))
```

```{r stability-comparison, eval=has_dsen_bootstrap, fig.height=5}
# Compare mean selection frequency
stability_long <- stability_analysis %>%
  select(drug, en_mean_freq, dsen_mean_freq) %>%
  pivot_longer(cols = c(en_mean_freq, dsen_mean_freq),
               names_to = "method", values_to = "mean_freq") %>%
  mutate(method = ifelse(method == "en_mean_freq", "Standard EN", "DSEN"))

p_stability <- ggplot(stability_long, aes(x = method, y = mean_freq, fill = method)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Standard EN" = "#4DAF4A", "DSEN" = "#FF7F00")) +
  labs(
    title = "Mean Selection Frequency by Method",
    subtitle = "Higher frequency = more stable feature selection",
    x = NULL,
    y = "Mean Bootstrap Selection Frequency"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "none"
  )

print(p_stability)

ggsave(file.path(fig_dir, "feature_stability_en_vs_dsen.pdf"),
       p_stability, width = 6, height = 5)

# T-test for stability difference
t_result <- t.test(stability_analysis$dsen_mean_freq, stability_analysis$en_mean_freq,
                   paired = TRUE)
cat(sprintf("\nPaired t-test (DSEN vs EN mean frequency): t = %.2f, p = %.2e\n",
            t_result$statistic, t_result$p.value))
```

## Stable vs Unstable Features

```{r stable-unstable, eval=has_dsen_bootstrap, fig.height=5}
# Compare stable feature counts
stable_comparison <- stability_analysis %>%
  select(drug, en_n_stable, dsen_n_stable) %>%
  pivot_longer(cols = c(en_n_stable, dsen_n_stable),
               names_to = "method", values_to = "n_stable") %>%
  mutate(method = ifelse(method == "en_n_stable", "Standard EN", "DSEN"))

p_stable <- ggplot(stable_comparison, aes(x = method, y = n_stable, fill = method)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Standard EN" = "#4DAF4A", "DSEN" = "#FF7F00")) +
  labs(
    title = "Number of Highly Stable Features (freq >= 80%)",
    subtitle = "Features selected in 80% or more of bootstrap resamples",
    x = NULL,
    y = "Number of Stable Features"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "none"
  )

print(p_stable)

ggsave(file.path(fig_dir, "feature_stable_count.pdf"),
       p_stable, width = 6, height = 5)
```

## Interpretation

DSEN shows higher mean selection frequency than EN, indicating that when DSEN selects
a feature, it does so more consistently across bootstrap resamples. This increased
stability likely results from the structured penalty that enforces consistency
between shared and tissue-specific effects.

# Part 4: Shared vs Tissue-Specific Features in DSEN

## Rationale

A unique property of DSEN is its ability to distinguish between:

- **Shared features**: Non-zero coefficient in the shared component (affects all tissues)
- **Tissue-specific features**: Non-zero coefficient in tissue-specific components

This decomposition provides biological interpretability: shared features represent
pan-cancer predictors, while tissue-specific features represent context-dependent effects.

```{r shared-tissue-analysis, eval=has_dsen_bootstrap}
# Analyze the shared vs tissue-specific decomposition
coef_analysis <- map_dfr(dsen_bootstrap_results, function(r) {
  if (is.null(r$dsen_bootstrap$mean_coefs)) return(NULL)

  mean_coefs <- r$dsen_bootstrap$mean_coefs

  # Shared column is first
  shared_nonzero <- sum(mean_coefs[, "Shared"] != 0)

  # Tissue columns are the rest
  tissue_cols <- colnames(mean_coefs)[colnames(mean_coefs) != "Shared"]
  tissue_coefs <- mean_coefs[, tissue_cols, drop = FALSE]

  # Features with any tissue-specific effect
  tissue_specific_any <- sum(rowSums(tissue_coefs != 0) > 0)

  # Features with ONLY shared effect
  only_shared <- sum(mean_coefs[, "Shared"] != 0 &
                     rowSums(tissue_coefs != 0) == 0)

  # Features with ONLY tissue-specific effect (no shared)
  only_tissue <- sum(mean_coefs[, "Shared"] == 0 &
                     rowSums(tissue_coefs != 0) > 0)

  # Features with both
  both <- sum(mean_coefs[, "Shared"] != 0 &
              rowSums(tissue_coefs != 0) > 0)

  tibble(
    drug = r$drug,
    n_shared = shared_nonzero,
    n_tissue_specific = tissue_specific_any,
    n_only_shared = only_shared,
    n_only_tissue = only_tissue,
    n_both = both,
    n_total = shared_nonzero + only_tissue
  )
})

cat("\n=== Shared vs Tissue-Specific Features ===\n")
cat(sprintf("Mean shared features: %.1f\n", mean(coef_analysis$n_shared)))
cat(sprintf("Mean tissue-specific features: %.1f\n", mean(coef_analysis$n_tissue_specific)))
cat(sprintf("Mean features with ONLY shared effect: %.1f\n", mean(coef_analysis$n_only_shared)))
cat(sprintf("Mean features with ONLY tissue effect: %.1f\n", mean(coef_analysis$n_only_tissue)))
cat(sprintf("Mean features with BOTH effects: %.1f\n", mean(coef_analysis$n_both)))

# Drugs with no shared features
no_shared <- sum(coef_analysis$n_shared == 0)
cat(sprintf("\nDrugs with 0 shared features: %d (%.1f%%)\n",
            no_shared, 100 * no_shared / nrow(coef_analysis)))
```

```{r shared-tissue-scatter, eval=has_dsen_bootstrap, fig.height=6}
p_shared_tissue <- ggplot(coef_analysis, aes(x = n_only_shared, y = n_only_tissue)) +
  geom_point(alpha = 0.6, size = 2, color = "#666666") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  labs(
    title = "Shared-Only vs Tissue-Only Features in DSEN",
    subtitle = "Features classified by whether they have shared effects, tissue effects, or both",
    x = "Features with ONLY Shared Effect",
    y = "Features with ONLY Tissue-Specific Effect"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank())

print(p_shared_tissue)

ggsave(file.path(fig_dir, "feature_shared_vs_tissue.pdf"),
       p_shared_tissue, width = 8, height = 6)
```

```{r shared-distribution, eval=has_dsen_bootstrap, fig.height=5}
p_shared_dist <- ggplot(coef_analysis, aes(x = n_shared)) +
  geom_histogram(bins = 30, fill = "#FF7F00", color = "white", alpha = 0.8) +
  geom_vline(xintercept = mean(coef_analysis$n_shared),
             color = "#984EA3", linewidth = 1) +
  annotate("text", x = mean(coef_analysis$n_shared) + 5, y = Inf,
           label = sprintf("Mean: %.0f", mean(coef_analysis$n_shared)),
           vjust = 2, hjust = 0, color = "#984EA3", size = 4) +
  annotate("text", x = 5, y = 30,
           label = sprintf("%d drugs (%.1f%%)\nhave 0 shared features",
                          no_shared, 100 * no_shared / nrow(coef_analysis)),
           hjust = 0, color = "#E41A1C", size = 3.5) +
  labs(
    title = "Distribution of Shared Features per Drug",
    x = "Number of Shared Features",
    y = "Number of Drugs"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank())

print(p_shared_dist)

ggsave(file.path(fig_dir, "feature_shared_distribution.pdf"),
       p_shared_dist, width = 8, height = 5)
```

## Interpretation

The split between shared and tissue-specific features varies widely across drugs.
Some drugs are driven almost entirely by pan-cancer features, while others have
substantial tissue-specific components.

Drugs with 0 shared features (8.3%) are particularly interesting - their response
is entirely tissue-dependent, suggesting highly specific mechanisms that may be
worth investigating biologically.

# Summary Tables

## Feature Stability Summary

```{r summary-table, eval=has_dsen_bootstrap}
summary_df <- tibble(
  Metric = c(
    "Mean EN features selected",
    "Mean DSEN features selected",
    "Feature reduction (DSEN vs EN)",
    "DSEN more sparse (%)",
    "---",
    "Mean EN selection frequency",
    "Mean DSEN selection frequency",
    "---",
    "Mean EN stable features (>80%)",
    "Mean DSEN stable features (>80%)",
    "---",
    "Mean shared features",
    "Mean tissue-specific features",
    "Drugs with 0 shared features"
  ),
  Value = c(
    sprintf("%.0f", mean(dsen_feature_df$n_en)),
    sprintf("%.0f", mean(dsen_feature_df$n_dsen)),
    sprintf("%.0f%%", mean(dsen_feature_df$reduction_pct)),
    sprintf("%.1f%%", 100 * mean(dsen_feature_df$dsen_more_sparse)),
    "",
    sprintf("%.3f", mean(stability_analysis$en_mean_freq)),
    sprintf("%.3f", mean(stability_analysis$dsen_mean_freq)),
    "",
    sprintf("%.1f", mean(stability_analysis$en_n_stable)),
    sprintf("%.1f", mean(stability_analysis$dsen_n_stable)),
    "",
    sprintf("%.0f", mean(coef_analysis$n_shared)),
    sprintf("%.0f", mean(coef_analysis$n_tissue_specific)),
    sprintf("%d (%.1f%%)", no_shared, 100 * no_shared / nrow(coef_analysis))
  )
)

kable(summary_df, caption = "Feature Selection Summary: EN vs DSEN")

write_csv(summary_df, file.path(table_dir, "feature_stability_summary.csv"))
```

# Conclusions

## Key Findings

1. **Data leakage changes feature selection**: Wrong CV selects substantially different
   features than correct CV (mean Jaccard ~ 0.2-0.3). This has major implications for
   biomarker discovery and validation.

2. **DSEN provides sparser models**: DSEN selects 42% fewer features than standard EN,
   with the structured penalty encouraging parsimony.

3. **DSEN improves selection stability**: Features selected by DSEN have higher bootstrap
   frequencies, indicating more reliable associations with drug response.

4. **Shared vs tissue-specific decomposition**: DSEN provides biological interpretability
   by distinguishing pan-cancer predictors from tissue-specific effects. 8.3% of drugs
   have purely tissue-specific predictors.

## Implications

- **For biomarker discovery**: Use correct CV with per-fold feature selection to avoid
  false discoveries. Consider DSEN for more stable, interpretable feature sets.

- **For model interpretation**: DSEN's shared/tissue-specific decomposition can guide
  biological investigation of drug mechanisms.

- **For validation**: Features selected stably across bootstrap resamples (high frequency)
  are better candidates for external validation.

# Session Info

```{r session-info}
sessionInfo()
```
