---
title: "Cross-Tissue Prediction Analysis"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# Executive Summary

## Background

This analysis tests whether drug response models trained on one tissue type can
generalize to predict response in other tissues. This is a critical question for:

1. **Clinical translation**: Can we use models trained on available data to predict
   response in rare cancer types?

2. **Biological insight**: Are drug response mechanisms tissue-specific or universal?

3. **Data leakage detection**: Cross-tissue prediction provides an independent test
   that's robust to within-tissue data leakage.

# Setup

```{r load-packages}
library(here)
library(tidyverse)
library(glmnet)
```

```{r load-methods}
source(here("methods", "00_paths.R"))
source(here("methods", "01_data_utils.R"))
source(here("methods", "02_cv_utils.R"))
source(here("methods", "04_plot_utils.R"))
```

# Configuration

```{r config}
MIN_SAMPLES_PER_TISSUE <- 20
GLOBAL_SEED <- 42

results_dir <- file.path(paths$results, "cross_tissue")
if (!dir.exists(results_dir)) dir.create(results_dir, recursive = TRUE)
```

# Load Data

```{r load-data}
data <- load_main_data("Sanger")
print_data_summary(data)
```

# Tissue Distribution

```{r tissue-dist}
tissue_counts <- table(data$response$tissue)
tissue_counts <- sort(tissue_counts, decreasing = TRUE)

# Filter to tissues with enough samples
large_tissues <- names(tissue_counts)[tissue_counts >= MIN_SAMPLES_PER_TISSUE]

cat(sprintf("Tissues with >= %d samples: %d\n", MIN_SAMPLES_PER_TISSUE, length(large_tissues)))
print(tissue_counts[large_tissues])
```

# Load Precomputed Results

```{r load-precomputed-results}
results_file <- file.path(paths$scratch, "cross_tissue_results.Rda")

if (!file.exists(results_file)) {
  stop(
    "Pre-computed results not found.\n",
    "Run: Rscript methods/run_cross_tissue.R\n",
    "Or:  TEST_MODE=TRUE Rscript methods/run_cross_tissue.R"
  )
}

load(results_file)

if (!exists("cross_tissue_results")) {
  stop("Expected object 'cross_tissue_results' in results file.")
}

cross_results_df <- cross_tissue_results %>% filter(status == "success")

cat("Loaded pre-computed cross-tissue results\n")
cat("  Rows:", nrow(cross_results_df), "\n")
cat("  Drugs:", length(unique(cross_results_df$drug)), "\n\n")
```

# Cross-Tissue Prediction Function

```{r cross-tissue-function}
#' Perform cross-tissue prediction
#'
#' @param data Data from load_main_data()
#' @param drug Drug name
#' @param train_tissue Tissue to train on
#' @param test_tissues Tissues to test on (vector)
#' @param alpha Elastic net alpha
#' @return Data frame with results
cross_tissue_predict <- function(data, drug, train_tissue, test_tissues,
                                  alpha = 0.5, cor_thresh = 0.1) {

  # Prepare full drug data
  drug_data <- prepare_drug_data(data, drug, min_samples = 10)
  if (is.null(drug_data)) return(NULL)

  # Split by tissue
  train_idx <- drug_data$tissue == train_tissue
  if (sum(train_idx) < 20) return(NULL)

  x_train <- drug_data$x[train_idx, ]
  y_train <- drug_data$y[train_idx]

  # Preprocess training data (feature selection + scaling)
  preproc <- preprocess_features(x_train, y_train,
                                  cor_thresh = cor_thresh,
                                  scale_features = TRUE)
  x_train_proc <- preproc$x

  if (ncol(x_train_proc) < 5) return(NULL)

  # Fit model on training tissue
  cv_fit <- cv.glmnet(x_train_proc, y_train, alpha = alpha, nfolds = 5)

  # Results storage
  results <- list()

  for (test_tissue in test_tissues) {
    test_idx <- drug_data$tissue == test_tissue
    if (sum(test_idx) < 5) next

    x_test <- drug_data$x[test_idx, ]
    y_test <- drug_data$y[test_idx]

    # Apply same preprocessing
    x_test_proc <- apply_preprocessing(x_test, preproc)

    # Ensure same features
    common_features <- intersect(colnames(x_train_proc), colnames(x_test_proc))
    if (length(common_features) < 5) next

    # Predict
    y_pred <- predict(cv_fit, newx = x_test_proc[, common_features],
                      s = "lambda.min")

    # Calculate metrics
    mse <- mean((y_test - y_pred)^2)
    cor_val <- cor(y_test, y_pred)
    r2 <- 1 - mse / var(y_test)

    results[[test_tissue]] <- tibble(
      drug = drug,
      train_tissue = train_tissue,
      test_tissue = test_tissue,
      n_train = sum(train_idx),
      n_test = sum(test_idx),
      mse = mse,
      correlation = cor_val,
      r2 = r2
    )
  }

  bind_rows(results)
}
```

# Example: Cross-Tissue for One Drug

```{r example-drug}
# Pick a drug with good sample size across tissues
example_drug <- "Erlotinib"

# Pick a well-represented tissue for training
train_tissue <- "lung_NSCLC"
test_tissues <- setdiff(large_tissues, train_tissue)

cat(sprintf("Drug: %s\n", example_drug))
cat(sprintf("Training tissue: %s\n", train_tissue))
cat(sprintf("Test tissues: %s\n", paste(test_tissues, collapse = ", ")))

# Run cross-tissue prediction
example_results <- cross_results_df %>%
  filter(drug == example_drug, strategy == "cross") %>%
  arrange(desc(r2))

if (nrow(example_results) > 0) {
  knitr::kable(example_results, digits = 3,
               caption = sprintf("Cross-Tissue Prediction: %s (train all other tissues)",
                                 example_drug))
}
```

# Within-Tissue vs Cross-Tissue Performance

```{r within-vs-cross}
# Compare within-tissue CV to cross-tissue prediction (precomputed)
within_row <- cross_results_df %>%
  filter(drug == example_drug, strategy == "within", train_tissue == train_tissue)

cross_row <- cross_results_df %>%
  filter(drug == example_drug, strategy == "cross", test_tissue == train_tissue)

reverse_row <- cross_results_df %>%
  filter(drug == example_drug, strategy == "reverse", train_tissue == train_tissue)

if (nrow(within_row) > 0) {
  cat(sprintf("Within-tissue CV MSE (%s): %.4f\n",
              train_tissue, within_row$mse[1]))
}

if (nrow(cross_row) > 0) {
  cat(sprintf("Cross-tissue MSE (train all other -> test %s): %.4f\n",
              train_tissue, cross_row$mse[1]))
}

if (nrow(reverse_row) > 0) {
  cat(sprintf("Reverse MSE (train %s -> test all other): %.4f\n",
              train_tissue, reverse_row$mse[1]))
}
```

# Full Cross-Tissue Analysis

```{r full-analysis}
# Results are loaded from precomputed scratch output.
```

# Visualize Cross-Tissue Performance

```{r visualize-cross, fig.height=7}
if (nrow(cross_results_df) > 0) {
  ggplot(cross_results_df, aes(x = test_tissue, y = r2, fill = strategy)) +
    geom_boxplot(alpha = 0.7) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    facet_wrap(~ strategy, scales = "free_x") +
    labs(
      title = "Cross-Tissue Prediction Performance",
      subtitle = "R2 by test tissue, faceted by evaluation strategy",
      x = "Test Tissue",
      y = "R2"
    ) +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

## Summary by Drug

```{r summary-drug}
if (nrow(cross_results_df) > 0) {
  drug_summary <- cross_results_df %>%
    group_by(drug, strategy) %>%
    summarize(
      N_experiments = n(),
      Mean_R2 = mean(r2, na.rm = TRUE),
      Median_R2 = median(r2, na.rm = TRUE),
      Pct_positive_R2 = 100 * mean(r2 > 0, na.rm = TRUE)
    ) %>%
    arrange(desc(Mean_R2))

  knitr::kable(drug_summary, digits = 3,
               caption = "Cross-Tissue Performance by Drug")
}
```

# Save Results

```{r save-results}
if (nrow(cross_results_df) > 0) {
  write_csv(cross_results_df, file.path(results_dir, "cross_tissue_results.csv"))
  cat(sprintf("Results saved to: %s\n", results_dir))
}
```

# Conclusions

1. **Cross-tissue generalization varies**: Some drugs show reasonable cross-tissue
   prediction (positive R²) while others fail completely.

2. **Implications for clinical translation**: Models should be validated across
   tissue types before being applied to new cancer types.

3. **Tissue-specific effects**: Poor cross-tissue generalization suggests that
   drug response mechanisms may be tissue-dependent, supporting the use of
   tissue-aware models like DSEN.

# Session Info

```{r session-info}
sessionInfo()
```
