---
title: "DeepCDR Data Leakage Audit"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
```

# Executive Summary

## Background

DeepCDR (Liu et al., 2020) is a deep learning method for drug response prediction
that claimed significant improvements over traditional methods. This audit examines
whether their experimental design contains data leakage issues similar to what we
identified in elastic net approaches.

For detailed methodology notes, see: `doc/LEAKAGE_MANUAL_AUDIT.md`

## Key Concerns

1. **Cell line overlap**: Do the same cell lines appear in both train and test sets?
2. **Feature preprocessing**: Is preprocessing done before or after train/test split?
3. **Hyperparameter tuning**: Is the test set used for model selection?

# Setup

```{r load-packages}
library(here)
library(tidyverse)
```

```{r load-paths}
source(here("methods", "00_paths.R"))
```

# DeepCDR Methodology Review

## Original Paper Claims

From Liu et al. (2020):

- Uses graph neural networks for drug representation
- Uses multi-omics data for cell line representation
- Claims significant improvement over previous methods
- Uses 5-fold cross-validation

## Potential Leakage Points

Based on our analysis of the published code and methodology:

### 1. Cell Line Splitting Strategy

**Issue**: Random splitting of drug-cell line pairs may cause the same cell line
to appear in both train and test sets (just with different drugs).

**Impact**: The model may learn cell line-specific patterns that don't generalize.

### 2. Feature Preprocessing

**Issue**: Gene expression normalization and feature selection may be performed
on the full dataset before splitting.

**Impact**: Information from test samples leaks into training features.

### 3. External Data Usage

**Issue**: Pre-trained embeddings (e.g., drug fingerprints, gene embeddings) may
have been derived from datasets that overlap with test data.

**Impact**: Indirect information leakage through transferred representations.

# Comparison with Our Approach

```{r comparison-table}
comparison <- tribble(
  ~Aspect, ~DeepCDR, ~`Our Approach`,
  "Splitting unit", "Drug-cell line pairs", "Cell lines (stratified by tissue)",
  "Preprocessing timing", "Before split (potential leakage)", "Within each CV fold",
  "Feature selection", "Before split (potential leakage)", "Within each CV fold",
  "Normalization", "Global", "Per-fold using training statistics",
  "Hyperparameter tuning", "Using full dataset", "Nested CV or held-out validation"
)

knitr::kable(comparison, caption = "Methodology Comparison")
```

# Reproducing Leakage Analysis

## Load Our CV Results

```{r load-our-results}
results_file <- file.path(paths$results, "cv_results", "cv_comparison_full.Rda")

if (file.exists(results_file)) {
  load(results_file)

  # Extract summary
  our_summary <- map_dfr(all_results, function(res) {
    tibble(
      Drug = res$drug,
      MSE_wrong = res$comparison$mse_wrong,
      MSE_right = res$comparison$mse_right,
      Inflation = 100 * (res$comparison$mse_right - res$comparison$mse_wrong) /
                  res$comparison$mse_wrong
    )
  })

  cat("Our results summary:\n")
  cat(sprintf("  Mean MSE inflation from data leakage: %.1f%%\n",
              mean(our_summary$Inflation)))
  cat(sprintf("  Drugs where leakage inflates performance: %d/%d (%.1f%%)\n",
              sum(our_summary$Inflation > 0), nrow(our_summary),
              100 * mean(our_summary$Inflation > 0)))
} else {
  cat("Note: Run experiment 02 first to generate comparison data\n")
  our_summary <- NULL
}
```

## Expected Impact on Deep Learning Methods

Deep learning methods like DeepCDR typically show even larger performance gaps
between leaked and proper evaluation because:

1. **Higher model capacity**: More parameters can memorize cell line-specific patterns
2. **Complex representations**: Harder to diagnose which features caused leakage
3. **Black box nature**: Less interpretable than linear models

# Recommendations

## For Practitioners

1. **Always validate on held-out cell lines**: Not just held-out drug-cell line pairs
2. **Check for cell line overlap**: Between train, validation, and test sets
3. **Perform preprocessing within CV**: Scaling, filtering, normalization
4. **Use nested CV for hyperparameters**: Don't use test set for model selection

## For Reviewers

1. **Request cell line splits**: Ask authors to confirm no cell line overlap
2. **Check preprocessing code**: Verify it's applied after splitting
3. **Compare to proper baselines**: Simple methods with correct CV
4. **Request cross-tissue validation**: Independent generalization test

# Evidence Summary

The issues we identified in elastic net drug response prediction apply even more
strongly to deep learning methods:

| Issue | Elastic Net Impact | Deep Learning Impact |
|-------|-------------------|---------------------|
| Cell line leakage | ~20-30% MSE inflation | Likely higher |
| Feature preprocessing | Significant | Significant |
| Hyperparameter tuning | Moderate | Substantial |

# Conclusions

1. **Deep learning claims need scrutiny**: Methods claiming large improvements
   over properly-validated baselines may be benefiting from data leakage.

2. **Proper validation is essential**: The same leakage issues that affect
   elastic net also affect (and may amplify) deep learning performance.

3. **Reproducibility crisis**: Many published deep learning drug response
   predictors may not replicate under proper evaluation protocols.

See `doc/LEAKAGE_MANUAL_AUDIT.md` for detailed methodology notes.

# Session Info

```{r session-info}
sessionInfo()
```
