---
title: "Publication Figures for Nature Methods"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

<!--
  To reproduce DSEN results from scratch:
  1. Run methods/run_dsen_production.R with MIN_SAMPLES_PER_TISSUE=25
  2. Run methods/run_dsen_bootstrap.R
  3. Run experiments/11_tissue_target_recovery.Rmd

  To regenerate just the publication figures, run:
    Rscript experiments/generate_v2_figures.R

  This Rmd notebook loads from results/dsen_results/threshold_25_v2/ and
  results/tables/threshold_25_v2/ (the v2 DSEN results with bug fixes).
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 300
)
```

# Executive Summary

This notebook generates publication-quality figures for the manuscript submitted to
Nature Communications. The paper has three main contributions:

1. **Data Leakage Identification**: Demonstrating that published CV implementations
   contain data leakage (feature selection before train/test split)

2. **Correct Implementation**: Providing proper CV with per-fold feature selection
   and correct SE calculation (not SD)

3. **DSEN Method**: Tissue-aware elastic net that improves prediction for 93% of drugs

All figures are saved in both PDF (for publication) and PNG (for preview) formats
in `results/figures/`.

# Setup

```{r load-packages}
library(here)
library(tidyverse)
library(ggpubr)
library(patchwork)
library(scales)
library(viridis)
library(RColorBrewer)
library(knitr)
```

```{r load-paths}
source(here("methods", "00_paths.R"))
source(here("methods", "04_plot_utils.R"))
```

```{r output-dirs}
fig_dir <- file.path(paths$results, "figures")
table_dir <- file.path(paths$results, "tables")
for (d in c(fig_dir, table_dir)) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}
```

# Color Scheme

```{r color-scheme}
# Define consistent publication color scheme
# Colorblind-friendly palette based on ColorBrewer
colors <- list(
  # Main comparisons
  wrong = "#E41A1C",       # Red - incorrect/leakage
  right = "#377EB8",       # Blue - correct
  dsen = "#FF7F00",        # Orange - DSEN
  en = "#4DAF4A",          # Green - standard EN

  # Secondary
  neutral = "#666666",     # Gray
  highlight = "#984EA3",   # Purple

  # Gradients
  diverging_low = "#2166AC",
  diverging_mid = "#F7F7F7",
  diverging_high = "#B2182B"
)

# Custom theme for publication
theme_pub <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      panel.grid.minor = element_blank(),
      panel.border = element_rect(color = "gray80", fill = NA, linewidth = 0.5),
      axis.text = element_text(color = "black"),
      axis.ticks = element_line(color = "gray80"),
      legend.position = "bottom",
      plot.title = element_text(face = "bold", hjust = 0),
      plot.subtitle = element_text(color = "gray40"),
      strip.background = element_rect(fill = "gray95"),
      strip.text = element_text(face = "bold")
    )
}
```

# Load All Pre-computed Results

```{r load-results}
# DSEN Production Results (v2: threshold_25_v2 with bug fixes)
dsen_file <- here("results", "tables", "threshold_25_v2", "dsen_full_results.csv")
if (file.exists(dsen_file)) {
  dsen_df <- read_csv(dsen_file, show_col_types = FALSE) %>%
    filter(status == "success")
  cat("Loaded DSEN results:", nrow(dsen_df), "drugs\n")
} else {
  stop("DSEN results not found. Run methods/run_dsen_production.R first.")
}

# DSEN Bootstrap Results (v2)
bootstrap_file <- here("results", "dsen_results", "threshold_25_v2", "dsen_bootstrap_results.Rda")
if (file.exists(bootstrap_file)) {
  load(bootstrap_file)
  bootstrap_results <- all_results
  rm(all_results)
  cat("Loaded bootstrap results:", length(bootstrap_results), "drugs\n")
} else {
  bootstrap_results <- NULL
  cat("Bootstrap results not found - some figures will be limited\n")
}

# Try to load CV wrong vs right results (if available)
cv_files <- c(
  file.path(paths$scratch, "cv_wrong_vs_right_results.Rda"),
  here("results", "cv_results", "cv_comparison_full.Rda")
)
cv_file <- cv_files[file.exists(cv_files)][1]

cv_df <- NULL
if (!is.na(cv_file)) {
  tryCatch({
    load(cv_file)
    if (exists("cv_results")) {
      cv_df <- cv_results %>% filter(status == "success")
    } else if (exists("all_results")) {
      cv_df <- map_dfr(all_results, function(res) {
        tryCatch({
          tibble(
            drug = res$drug,
            n_samples = res$n_samples,
            mse_wrong = res$comparison$mse_wrong,
            mse_right = res$comparison$mse_right,
            mse_pct_increase = res$comparison$mse_pct_increase,
            jaccard = res$comparison$jaccard
          )
        }, error = function(e) NULL)
      })
      rm(all_results)
    }
    # Validate cv_df has required columns
    if (!is.null(cv_df) && nrow(cv_df) > 0) {
      required_cols <- c("mse_wrong", "mse_right", "mse_pct_increase")
      if (!all(required_cols %in% names(cv_df))) {
        cv_df <- NULL
      }
    }
    if (!is.null(cv_df)) cat("Loaded CV comparison results:", nrow(cv_df), "drugs\n")
  }, error = function(e) {
    cv_df <<- NULL
    cat("Error loading CV results:", e$message, "\n")
  })
}

if (is.null(cv_df)) {
  cat("CV comparison results not found or invalid - leakage figures will be skipped\n")
}
```

# Figure 1: DSEN Performance Overview (Main Figure)

## Rationale

Figure 1 is the main result figure showing DSEN improvement over standard EN.
We present a three-panel figure: (A) sorted bar chart showing per-drug improvement,
(B) distribution of improvements, (C) scatter of EN MSE vs DSEN MSE.

```{r fig1-prep}
# Prepare data
plot_df <- dsen_df %>%
  arrange(pct_improvement) %>%
  mutate(
    rank = row_number(),
    direction = ifelse(pct_improvement > 0, "DSEN Better", "EN Better")
  )

# Statistics for annotation
n_dsen_better <- sum(dsen_df$pct_improvement > 0)
n_total <- nrow(dsen_df)
pct_dsen_wins <- 100 * n_dsen_better / n_total
mean_improvement <- mean(dsen_df$pct_improvement)
median_improvement <- median(dsen_df$pct_improvement)
```

## Panel A: DSEN Improvement Waterfall

```{r fig1a, fig.width=10, fig.height=5}
fig1a <- ggplot(plot_df, aes(x = rank, y = pct_improvement, fill = direction)) +
  geom_col(width = 1, color = NA) +
  scale_fill_manual(
    values = c("DSEN Better" = colors$dsen, "EN Better" = colors$en),
    name = NULL
  ) +
  geom_hline(yintercept = 0, color = "gray20", linewidth = 0.5) +
  geom_hline(yintercept = mean_improvement,
             linetype = "dashed", color = colors$highlight, linewidth = 0.7) +
  annotate("text", x = 20, y = mean_improvement + 1.5,
           label = sprintf("Mean: %.1f%%", mean_improvement),
           color = colors$highlight, hjust = 0, size = 3.5) +
  labs(
    x = "Drug (ranked by improvement)",
    y = "MSE Improvement (%)",
    title = NULL
  ) +
  scale_x_continuous(expand = c(0.01, 0)) +
  theme_pub() +
  theme(legend.position = c(0.85, 0.15))

print(fig1a)
```

## Panel B: Improvement Distribution

```{r fig1b, fig.width=5, fig.height=5}
fig1b <- ggplot(dsen_df, aes(x = pct_improvement)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 35, fill = colors$dsen, color = "white", alpha = 0.8) +
  geom_density(color = "darkorange4", linewidth = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray30", linewidth = 0.7) +
  geom_vline(xintercept = mean_improvement, color = colors$highlight, linewidth = 0.7) +
  annotate("text", x = mean_improvement + 1, y = 0.11,
           label = sprintf("Mean: %.1f%%", mean_improvement),
           color = colors$highlight, hjust = 0, size = 3.5) +
  annotate("text", x = -8, y = 0.11,
           label = sprintf("DSEN wins: %.0f%%", pct_dsen_wins),
           color = colors$dsen, hjust = 0, size = 3.5, fontface = "bold") +
  labs(
    x = "MSE Improvement (%)",
    y = "Density",
    title = NULL
  ) +
  scale_x_continuous(limits = c(-12, 25)) +
  theme_pub() +
  theme(legend.position = "none")

print(fig1b)
```

## Panel C: EN vs DSEN Scatter

```{r fig1c, fig.width=5, fig.height=5}
fig1c <- ggplot(dsen_df, aes(x = mse_en, y = mse_dsen)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = pct_improvement > 0), alpha = 0.6, size = 1.5) +
  scale_color_manual(
    values = c("TRUE" = colors$dsen, "FALSE" = colors$en),
    labels = c("TRUE" = "DSEN Better", "FALSE" = "EN Better"),
    name = NULL
  ) +
  labs(
    x = "MSE (Standard EN)",
    y = "MSE (DSEN)",
    title = NULL
  ) +
  coord_fixed(xlim = c(0, 0.12), ylim = c(0, 0.12)) +
  theme_pub() +
  theme(legend.position = c(0.75, 0.15))

print(fig1c)
```

## Combined Figure 1

```{r fig1-combined, fig.width=14, fig.height=5}
fig1_combined <- fig1a + fig1b + fig1c +
  plot_annotation(
    tag_levels = "A",
    title = "DSEN Improves Drug Response Prediction",
    subtitle = sprintf("Analysis of %d drugs | DSEN outperforms standard EN for %.0f%% of drugs",
                       n_total, pct_dsen_wins),
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(color = "gray40", size = 11)
    )
  ) &
  theme(plot.tag = element_text(face = "bold", size = 14))

print(fig1_combined)

# Save
ggsave(file.path(fig_dir, "fig1_dsen_overview.pdf"),
       fig1_combined, width = 14, height = 5)
ggsave(file.path(fig_dir, "fig1_dsen_overview.png"),
       fig1_combined, width = 14, height = 5, dpi = 300)
```

# Figure 2: Top Performers and Biological Context

## Rationale

Highlight drugs where DSEN shows the largest improvement. These drugs likely have
strong tissue-specific biology, validating the DSEN approach.

```{r fig2, fig.width=10, fig.height=7}
# Top 15 drugs
top_drugs <- dsen_df %>%
  arrange(desc(pct_improvement)) %>%
  head(15) %>%
  mutate(drug = fct_reorder(drug, pct_improvement))

fig2 <- ggplot(top_drugs, aes(x = drug, y = pct_improvement)) +
  geom_col(fill = colors$dsen, alpha = 0.85) +
  geom_text(aes(label = sprintf("%.1f%%", pct_improvement)),
            hjust = -0.1, size = 3.5) +
  coord_flip() +
  labs(
    title = "Top 15 Drugs with Greatest DSEN Improvement",
    subtitle = "Drugs where tissue-specific modeling provides largest benefit over standard EN",
    x = NULL,
    y = "MSE Improvement (%)"
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  theme_pub(base_size = 12)

print(fig2)

ggsave(file.path(fig_dir, "fig2_top_performers.pdf"),
       fig2, width = 10, height = 7)
ggsave(file.path(fig_dir, "fig2_top_performers.png"),
       fig2, width = 10, height = 7, dpi = 300)
```

## Top Performers Table

```{r top-table}
top_table <- dsen_df %>%
  arrange(desc(pct_improvement)) %>%
  head(15) %>%
  select(drug, n_samples, n_tissues, mse_en, mse_dsen, pct_improvement)

kable(top_table, digits = 3,
      col.names = c("Drug", "Samples", "Tissues", "MSE (EN)", "MSE (DSEN)", "Improvement (%)"),
      caption = "Top 15 Drugs with Greatest DSEN Improvement")

write_csv(top_table, file.path(table_dir, "top_dsen_performers.csv"))
```

# Figure 3: DSEN vs Drug Difficulty

## Rationale

Examine whether DSEN provides more benefit for "hard" drugs (high baseline MSE).

```{r fig3, fig.width=10, fig.height=5}
fig3 <- ggplot(dsen_df, aes(x = mse_en, y = pct_improvement)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(alpha = 0.5, size = 2, color = colors$neutral) +
  geom_smooth(method = "loess", se = TRUE, span = 0.8,
              color = colors$dsen, fill = colors$dsen, alpha = 0.2) +
  labs(
    title = "DSEN Improvement vs Drug Prediction Difficulty",
    subtitle = "Hard-to-predict drugs (high MSE) show similar benefit from tissue-aware modeling",
    x = "Baseline MSE (Standard EN)",
    y = "DSEN Improvement (%)"
  ) +
  theme_pub()

print(fig3)

ggsave(file.path(fig_dir, "fig3_improvement_vs_difficulty.pdf"),
       fig3, width = 10, height = 5)
```

# Figure 4: Feature Selection Comparison

## Rationale

DSEN provides sparser models. This figure shows the feature count reduction
and its consistency across drugs.

```{r fig4-prep}
if (!is.null(bootstrap_results)) {
  feature_df <- map_dfr(bootstrap_results, function(r) {
    tibble(
      drug = r$drug,
      n_en = sum(r$en_bootstrap$cnt > 0),
      n_dsen = sum(r$dsen_bootstrap$cnt > 0)
    )
  }) %>%
    mutate(
      reduction = n_en - n_dsen,
      reduction_pct = 100 * reduction / n_en
    )

  has_feature_data <- TRUE
} else {
  has_feature_data <- FALSE
}
```

```{r fig4, fig.width=12, fig.height=5, eval=has_feature_data}
# Panel A: EN vs DSEN feature count scatter
fig4a <- ggplot(feature_df, aes(x = n_en, y = n_dsen)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(alpha = 0.5, size = 2, color = colors$neutral) +
  geom_smooth(method = "lm", se = TRUE, color = colors$dsen, fill = colors$dsen, alpha = 0.2) +
  labs(
    x = "Features (Standard EN)",
    y = "Features (DSEN)",
    title = NULL
  ) +
  coord_fixed() +
  theme_pub()

# Panel B: Feature reduction distribution
fig4b <- ggplot(feature_df, aes(x = reduction_pct)) +
  geom_histogram(bins = 35, fill = colors$dsen, color = "white", alpha = 0.8) +
  geom_vline(xintercept = mean(feature_df$reduction_pct),
             color = colors$highlight, linewidth = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray30") +
  annotate("text", x = mean(feature_df$reduction_pct) + 2, y = 25,
           label = sprintf("Mean: %.0f%%", mean(feature_df$reduction_pct)),
           color = colors$highlight, hjust = 0, size = 4) +
  labs(
    x = "Feature Count Reduction (%)",
    y = "Number of Drugs",
    title = NULL
  ) +
  theme_pub()

# Combined
fig4_combined <- fig4a + fig4b +
  plot_annotation(
    tag_levels = "A",
    title = "DSEN Provides Sparser Feature Selection",
    subtitle = sprintf("DSEN selects %.0f%% fewer features than standard EN on average",
                       mean(feature_df$reduction_pct)),
    theme = theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(color = "gray40", size = 11)
    )
  ) &
  theme(plot.tag = element_text(face = "bold", size = 14))

print(fig4_combined)

ggsave(file.path(fig_dir, "fig4_feature_sparsity.pdf"),
       fig4_combined, width = 12, height = 5)
ggsave(file.path(fig_dir, "fig4_feature_sparsity.png"),
       fig4_combined, width = 12, height = 5, dpi = 300)
```

# Figure 5: Lambda Ratio Distribution

## Rationale

The lambda ratio controls the balance between shared and tissue-specific coefficients.
This figure shows that most drugs prefer a ratio of 2 (moderate emphasis on shared).

```{r fig5, fig.width=8, fig.height=5}
lamrat_summary <- dsen_df %>%
  count(lamrat_dsen, name = "n_drugs") %>%
  mutate(pct = 100 * n_drugs / sum(n_drugs))

fig5 <- ggplot(dsen_df, aes(x = factor(lamrat_dsen))) +
  geom_bar(fill = colors$dsen, alpha = 0.85) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.3, size = 4) +
  labs(
    title = "Distribution of Optimal Lambda Ratios",
    subtitle = "Lambda ratio controls balance between shared and tissue-specific coefficients\n(Higher = stronger preference for shared features)",
    x = "Lambda Ratio",
    y = "Number of Drugs"
  ) +
  theme_pub()

print(fig5)

ggsave(file.path(fig_dir, "fig5_lamrat_distribution.pdf"),
       fig5, width = 8, height = 5)
```

# Supplementary Figures

## Figure S1: All Drugs Performance Table

```{r figs1-table}
# Full results table
full_table <- dsen_df %>%
  arrange(desc(pct_improvement)) %>%
  select(drug, n_samples, n_tissues, mse_en, mse_dsen, pct_improvement, lamrat_dsen)

kable(head(full_table, 30), digits = 3,
      caption = "DSEN Results Summary (Top 30 of 265 drugs)")

write_csv(full_table, file.path(table_dir, "dsen_full_results.csv"))
```

## Figure S2: Win Rate by Quartile

```{r figs2, fig.width=10, fig.height=5}
dsen_df_quartile <- dsen_df %>%
  mutate(
    difficulty_q = cut(mse_en,
                       breaks = quantile(mse_en, probs = c(0, 0.25, 0.5, 0.75, 1)),
                       labels = c("Q1 (Easy)", "Q2", "Q3", "Q4 (Hard)"),
                       include.lowest = TRUE),
    tissue_q = cut(n_tissues,
                   breaks = c(0, 27, 28, 29, 30),
                   labels = c("<27", "27-28", "29", "30"),
                   include.lowest = TRUE)
  )

# Win rate by difficulty
quartile_summary <- dsen_df_quartile %>%
  group_by(difficulty_q) %>%
  summarize(
    n_drugs = n(),
    win_rate = mean(pct_improvement > 0),
    mean_improvement = mean(pct_improvement),
    .groups = "drop"
  )

figS2a <- ggplot(quartile_summary, aes(x = difficulty_q, y = win_rate)) +
  geom_col(fill = colors$dsen, alpha = 0.85) +
  geom_text(aes(label = sprintf("%.0f%%\n(n=%d)", 100*win_rate, n_drugs)),
            vjust = -0.1, size = 3.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.85)) +
  labs(
    x = "Drug Difficulty Quartile",
    y = "DSEN Win Rate",
    title = "DSEN Win Rate by Drug Difficulty"
  ) +
  theme_pub()

# Win rate by tissue count
tissue_summary <- dsen_df_quartile %>%
  group_by(tissue_q) %>%
  summarize(
    n_drugs = n(),
    win_rate = mean(pct_improvement > 0),
    mean_improvement = mean(pct_improvement),
    .groups = "drop"
  )

figS2b <- ggplot(tissue_summary, aes(x = tissue_q, y = win_rate)) +
  geom_col(fill = colors$dsen, alpha = 0.85) +
  geom_text(aes(label = sprintf("%.0f%%\n(n=%d)", 100*win_rate, n_drugs)),
            vjust = -0.1, size = 3.5) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "gray50") +
  scale_y_continuous(labels = percent_format(), limits = c(0, 0.85)) +
  labs(
    x = "Number of Tissues",
    y = "DSEN Win Rate",
    title = "DSEN Win Rate by Tissue Count"
  ) +
  theme_pub()

figS2_combined <- figS2a + figS2b +
  plot_annotation(
    tag_levels = "A",
    theme = theme(plot.title = element_text(face = "bold"))
  )

print(figS2_combined)

ggsave(file.path(fig_dir, "figS2_winrate_by_quartile.pdf"),
       figS2_combined, width = 10, height = 5)
```

## Figure S3: Shared vs Tissue-Specific Features

```{r figs3, fig.width=10, fig.height=5, eval=has_feature_data}
# Analyze shared vs tissue coefficients
coef_df <- map_dfr(bootstrap_results, function(r) {
  if (is.null(r$dsen_bootstrap$mean_coefs)) return(NULL)

  mc <- r$dsen_bootstrap$mean_coefs
  shared_n <- sum(mc[, "Shared"] != 0)

  tissue_cols <- colnames(mc)[colnames(mc) != "Shared"]
  tissue_n <- sum(rowSums(mc[, tissue_cols, drop = FALSE] != 0) > 0)

  tibble(
    drug = r$drug,
    n_shared = shared_n,
    n_tissue_specific = tissue_n
  )
})

figS3a <- ggplot(coef_df, aes(x = n_shared, y = n_tissue_specific)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(alpha = 0.5, size = 2, color = colors$neutral) +
  labs(
    x = "Shared Features",
    y = "Tissue-Specific Features",
    title = "Shared vs Tissue-Specific Features in DSEN"
  ) +
  theme_pub()

# Drugs with 0 shared features
n_no_shared <- sum(coef_df$n_shared == 0)

figS3b <- ggplot(coef_df, aes(x = n_shared)) +
  geom_histogram(bins = 30, fill = colors$dsen, color = "white", alpha = 0.8) +
  geom_vline(xintercept = mean(coef_df$n_shared), color = colors$highlight, linewidth = 0.8) +
  annotate("text", x = 0, y = 35,
           label = sprintf("%d drugs (%.1f%%)\nhave 0 shared features",
                          n_no_shared, 100*n_no_shared/nrow(coef_df)),
           hjust = 0, color = colors$highlight, size = 3.5) +
  labs(
    x = "Number of Shared Features",
    y = "Number of Drugs",
    title = "Distribution of Shared Features per Drug"
  ) +
  theme_pub()

figS3_combined <- figS3a + figS3b +
  plot_annotation(tag_levels = "A")

print(figS3_combined)

ggsave(file.path(fig_dir, "figS3_shared_vs_tissue.pdf"),
       figS3_combined, width = 10, height = 5)
```

# CV Leakage Figures (If Available)

```{r cv-figures-check}
has_cv_data <- !is.null(cv_df) && nrow(cv_df) > 0
```

## Figure S4: CV Leakage Impact (Conditional)

```{r figs4, fig.width=12, fig.height=5, eval=has_cv_data}
# Panel A: MSE inflation distribution
figS4a <- ggplot(cv_df, aes(x = mse_pct_increase)) +
  geom_histogram(bins = 35, fill = colors$right, color = "white", alpha = 0.8) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray30") +
  geom_vline(xintercept = mean(cv_df$mse_pct_increase),
             color = colors$wrong, linewidth = 0.8) +
  annotate("text", x = mean(cv_df$mse_pct_increase) + 2, y = 20,
           label = sprintf("Mean: %.1f%%", mean(cv_df$mse_pct_increase)),
           color = colors$wrong, hjust = 0, size = 4) +
  labs(
    x = "MSE Increase (Correct vs Incorrect CV) (%)",
    y = "Number of Drugs",
    title = "Impact of CV Leakage on Performance Estimates"
  ) +
  theme_pub()

# Panel B: Wrong vs Right MSE scatter
figS4b <- ggplot(cv_df, aes(x = mse_wrong, y = mse_right)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
  geom_point(aes(color = mse_right > mse_wrong), alpha = 0.5, size = 2) +
  scale_color_manual(
    values = c("TRUE" = colors$wrong, "FALSE" = colors$right),
    labels = c("TRUE" = "Correct CV Higher (expected)", "FALSE" = "Incorrect CV Higher"),
    name = NULL
  ) +
  labs(
    x = "MSE (Incorrect CV - with leakage)",
    y = "MSE (Correct CV - no leakage)",
    title = "Comparison of MSE Estimates"
  ) +
  coord_fixed() +
  theme_pub() +
  theme(legend.position = c(0.7, 0.2))

figS4_combined <- figS4a + figS4b +
  plot_annotation(
    tag_levels = "A",
    title = "Data Leakage Inflates Apparent Performance",
    subtitle = sprintf("%.0f%% of drugs show inflated performance when using incorrect CV",
                       100 * mean(cv_df$mse_right > cv_df$mse_wrong))
  )

print(figS4_combined)

ggsave(file.path(fig_dir, "figS4_cv_leakage.pdf"),
       figS4_combined, width = 12, height = 5)
```

## Figure S5: Feature Stability (Jaccard) Distribution

```{r figs5, fig.width=8, fig.height=5, eval=has_cv_data}
if ("jaccard" %in% names(cv_df)) {
  figS5 <- ggplot(cv_df, aes(x = jaccard)) +
    geom_histogram(bins = 25, fill = colors$right, color = "white", alpha = 0.8) +
    geom_vline(xintercept = mean(cv_df$jaccard, na.rm = TRUE),
               color = colors$wrong, linewidth = 0.8) +
    annotate("text", x = mean(cv_df$jaccard, na.rm = TRUE) + 0.02, y = 15,
             label = sprintf("Mean: %.2f", mean(cv_df$jaccard, na.rm = TRUE)),
             color = colors$wrong, hjust = 0, size = 4) +
    annotate("text", x = 0.05, y = 20,
             label = sprintf("%.0f%% have\nJaccard = 0",
                            100 * mean(cv_df$jaccard == 0, na.rm = TRUE)),
             hjust = 0, color = colors$wrong, size = 3.5) +
    labs(
      title = "Feature Selection Overlap: Correct vs Incorrect CV",
      subtitle = "Low Jaccard indicates different features selected by each method",
      x = "Jaccard Similarity (Feature Overlap)",
      y = "Number of Drugs"
    ) +
    scale_x_continuous(limits = c(0, 1)) +
    theme_pub()

  print(figS5)

  ggsave(file.path(fig_dir, "figS5_jaccard_distribution.pdf"),
         figS5, width = 8, height = 5)
}
```

# Summary Statistics Table

```{r summary-table}
summary_stats <- tribble(
  ~Category, ~Metric, ~Value,
  "Dataset", "Total drugs analyzed", as.character(nrow(dsen_df)),
  "Dataset", "Mean samples per drug", sprintf("%.0f", mean(dsen_df$n_samples)),
  "Dataset", "Mean tissues per drug", sprintf("%.1f", mean(dsen_df$n_tissues)),
  "DSEN Performance", "DSEN win rate", sprintf("%.1f%%", pct_dsen_wins),
  "DSEN Performance", "Mean improvement", sprintf("%.2f%%", mean_improvement),
  "DSEN Performance", "Median improvement", sprintf("%.2f%%", median_improvement),
  "DSEN Performance", "Max improvement", sprintf("%.1f%% (Imatinib)", max(dsen_df$pct_improvement)),
  "DSEN Performance", "Min improvement", sprintf("%.1f%% (Crizotinib)", min(dsen_df$pct_improvement)),
  "Feature Selection", "Mean EN features", sprintf("%.0f", ifelse(has_feature_data, mean(feature_df$n_en), NA)),
  "Feature Selection", "Mean DSEN features", sprintf("%.0f", ifelse(has_feature_data, mean(feature_df$n_dsen), NA)),
  "Feature Selection", "Feature reduction", sprintf("%.0f%%", ifelse(has_feature_data, mean(feature_df$reduction_pct), NA)),
  "Parameters", "Alpha (all drugs)", "0.5",
  "Parameters", "Most common lambda ratio", "2"
)

kable(summary_stats, caption = "Summary Statistics for Publication")

write_csv(summary_stats, file.path(table_dir, "publication_summary_stats.csv"))
```

# Figure Manifest

```{r manifest}
manifest <- tribble(
  ~Figure, ~Filename, ~Description,
  "Figure 1", "fig1_dsen_overview.pdf", "Main DSEN result: waterfall, histogram, scatter",
  "Figure 2", "fig2_top_performers.pdf", "Top 15 drugs with greatest DSEN improvement",
  "Figure 3", "fig3_improvement_vs_difficulty.pdf", "DSEN improvement vs drug difficulty",
  "Figure 4", "fig4_feature_sparsity.pdf", "Feature selection comparison (sparsity)",
  "Figure 5", "fig5_lamrat_distribution.pdf", "Lambda ratio distribution",
  "Figure S2", "figS2_winrate_by_quartile.pdf", "Win rate by difficulty/tissue quartile",
  "Figure S3", "figS3_shared_vs_tissue.pdf", "Shared vs tissue-specific features",
  "Figure S4", "figS4_cv_leakage.pdf", "CV leakage impact (if data available)",
  "Figure S5", "figS5_jaccard_distribution.pdf", "Feature stability Jaccard"
)

kable(manifest, caption = "Figure Files Generated")

write_csv(manifest, file.path(fig_dir, "figure_manifest.csv"))

cat("\n=== All figures saved to ===\n")
cat(fig_dir, "\n")
```

# Session Info

```{r session-info}
sessionInfo()
```
