---
title: "Drug Target Validation: Do Models Find Known Targets?"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 150
)
```

# Executive Summary

## Objective

This analysis validates our drug response prediction models (EN and DSEN) by checking
whether they identify **known drug targets** as predictive features. If a model correctly
learns the biology, we expect drugs with known targets (e.g., EGFR for Erlotinib) to
show those targets among selected features.

## Key Questions

1. **Target Recovery Rate**: What proportion of drugs have their known target(s)
   among selected features?
2. **EN vs DSEN**: Does DSEN's tissue-aware approach improve or harm target recovery?
3. **Feature Type**: Do targets appear in expression, mutation, or copy number features?
4. **Biological Validation**: Does target recovery correlate with prediction performance?

# Setup

```{r load-packages}
library(here)
library(tidyverse)
library(knitr)
library(patchwork)
library(scales)
```

```{r load-paths}
source(here("methods", "00_paths.R"))
```

```{r output-dirs}
fig_dir <- file.path(paths$results, "figures")
table_dir <- file.path(paths$results, "tables")
```

# Load Data

## Drug Targets

```{r load-targets}
targets_file <- here("data", "drug_targets_curated.csv")

if (file.exists(targets_file)) {
  drug_targets <- read_csv(targets_file, show_col_types = FALSE)
  cat("Loaded drug targets:", nrow(drug_targets), "entries\n")
  cat("Unique drugs:", n_distinct(drug_targets$drug_name), "\n")
  cat("Unique target genes:", n_distinct(drug_targets$target_gene, na.rm = TRUE), "\n")
} else {
  stop("Drug targets file not found. Run: Rscript methods/process_drug_targets.R")
}

# Filter to drugs with known targets
targets_known <- drug_targets %>%
  filter(!is.na(target_gene)) %>%
  filter(source != "not_found")

cat("Drugs with at least one known target:",
    n_distinct(targets_known$drug_name), "\n")
```

## Gene Name Standardization

Many drug targets use common aliases (e.g., "P38" instead of MAPK14, "MEK1" instead
of MAP2K1). We map these to official HGNC gene symbols to enable proper matching
with our genomic features.

```{r load-gene-mapping}
mapping_file <- here("data", "gene_name_mapping.csv")

if (file.exists(mapping_file)) {
  gene_mapping <- read_csv(mapping_file, show_col_types = FALSE)
  cat("Loaded gene name mapping:", nrow(gene_mapping), "aliases\n")
} else {
  # Create empty mapping if file doesn't exist
  gene_mapping <- tibble(alias = character(), hgnc_symbol = character())
  cat("No gene mapping file found - using original names\n")
}

# Function to standardize gene names
standardize_gene <- function(gene, mapping_df) {
  # Check if gene is in mapping
  match_idx <- match(toupper(gene), toupper(mapping_df$alias))
  if (!is.na(match_idx)) {
    return(mapping_df$hgnc_symbol[match_idx])
  }
  return(gene)  # Return original if no mapping found
}

# Apply standardization to targets
targets_known <- targets_known %>%
  mutate(
    original_target = target_gene,
    target_gene = sapply(target_gene, standardize_gene, mapping_df = gene_mapping)
  )

# Show mapping examples
mapped_genes <- targets_known %>%
  filter(original_target != target_gene) %>%
  distinct(original_target, target_gene) %>%
  head(15)

cat("\n=== Gene Name Standardization Examples ===\n")
if (nrow(mapped_genes) > 0) {
  kable(mapped_genes, col.names = c("Original (alias)", "Standardized (HGNC)"))
}
```

## Feature Names from Data

```{r load-feature-names}
# Load the predictor matrix from CCLE data
load(file.path(paths$clean, "ccle.Rda"))

# Feature names from the predictor matrix x
# Format: GENE_mut.hsrec, GENE_Expr (for expression), etc.
feature_names <- colnames(x)

# Parse feature names to get gene and type
feature_info <- tibble(feature = feature_names) %>%
  mutate(
    # Determine feature type based on suffix patterns
    type = case_when(
      grepl("_mut\\.hsrec$|_mut$", feature) ~ "Mutation",
      grepl("_Expr$", feature) ~ "Expression",
      grepl("_CN$", feature) ~ "CopyNumber",
      grepl("_del$", feature) ~ "Deletion",
      grepl("_amp$", feature) ~ "Amplification",
      TRUE ~ "Unknown"
    ),
    # Extract gene name by removing suffix
    gene = sub("_mut\\.hsrec$|_mut$|_Expr$|_CN$|_del$|_amp$", "", feature)
  )

cat("Total features:", length(feature_names), "\n")
cat("Feature types:\n")
print(table(feature_info$type))

# Get unique genes in our feature set
genes_in_data <- unique(feature_info$gene)
cat("\nUnique genes in feature set:", length(genes_in_data), "\n")

# Clean up large objects
rm(x, y, cls)
```

## Check Target Coverage

```{r target-coverage}
# Check how many drug targets are actually present in our feature set
targets_in_features <- targets_known %>%
  mutate(
    target_in_features = target_gene %in% genes_in_data
  )

cat("=== Target Coverage in Feature Set ===\n")
coverage <- targets_in_features %>%
  group_by(drug_name) %>%
  summarize(
    n_targets = n(),
    n_covered = sum(target_in_features),
    any_covered = any(target_in_features)
  )

cat("Drugs with target in feature set:",
    sum(coverage$any_covered), "/",
    nrow(coverage), "\n")
cat("Percentage:", round(100 * mean(coverage$any_covered), 1), "%\n")

# Targets not in feature set
missing_targets <- targets_in_features %>%
  filter(!target_in_features) %>%
  distinct(target_gene)

cat("\nTargets NOT in our feature set (first 20):\n")
print(head(missing_targets, 20))
```

## Bootstrap Feature Selection Results

```{r load-bootstrap}
bootstrap_file <- here("results", "dsen_results", "dsen_bootstrap_results.Rda")

if (file.exists(bootstrap_file)) {
  load(bootstrap_file)
  bootstrap_results <- all_results
  rm(all_results)
  cat("Loaded bootstrap results for", length(bootstrap_results), "drugs\n")
} else {
  stop("Bootstrap results not found. Run: Rscript methods/run_dsen_bootstrap.R")
}
```

# Target Recovery Analysis

## Methodology

For each drug, we check if its known target gene(s) appear among the features
selected by EN and DSEN. A feature is considered "selected" if it was chosen
in at least 50% of bootstrap resamples (100 iterations).

```{r define-recovery-function}
# Selection frequency threshold
FREQ_THRESHOLD <- 0.5

# Function to check if target genes are in selected features
check_target_recovery <- function(drug_name, bootstrap_result, targets_df,
                                   feature_info_df, freq_threshold = 0.5) {

  # Get targets for this drug
  drug_targets <- targets_df %>%
    filter(drug_name == !!drug_name, target_in_features == TRUE) %>%
    pull(target_gene) %>%
    unique()

  if (length(drug_targets) == 0) {
    return(tibble(
      drug = drug_name,
      n_targets = 0,
      en_recovered = NA,
      dsen_recovered = NA,
      en_targets_found = NA_character_,
      dsen_targets_found = NA_character_
    ))
  }

  # EN: Get features selected >= threshold
  en_freq <- bootstrap_result$en_bootstrap$cnt / bootstrap_result$en_bootstrap$nboot
  en_selected_features <- names(en_freq)[en_freq >= freq_threshold]

  # Get genes from selected EN features
  en_genes <- feature_info_df %>%
    filter(feature %in% en_selected_features) %>%
    pull(gene) %>%
    unique()

  # DSEN: Get features selected >= threshold
  dsen_freq <- bootstrap_result$dsen_bootstrap$cnt / bootstrap_result$dsen_bootstrap$nboot
  dsen_selected_features <- names(dsen_freq)[dsen_freq >= freq_threshold]

  # Get genes from selected DSEN features
  dsen_genes <- feature_info_df %>%
    filter(feature %in% dsen_selected_features) %>%
    pull(gene) %>%
    unique()

  # Check recovery
  en_targets_found <- intersect(drug_targets, en_genes)
  dsen_targets_found <- intersect(drug_targets, dsen_genes)

  tibble(
    drug = drug_name,
    n_targets = length(drug_targets),
    en_n_selected = length(en_selected_features),
    dsen_n_selected = length(dsen_selected_features),
    en_n_genes = length(en_genes),
    dsen_n_genes = length(dsen_genes),
    en_recovered = length(en_targets_found) / length(drug_targets),
    dsen_recovered = length(dsen_targets_found) / length(drug_targets),
    en_n_found = length(en_targets_found),
    dsen_n_found = length(dsen_targets_found),
    en_targets_found = paste(en_targets_found, collapse = ";"),
    dsen_targets_found = paste(dsen_targets_found, collapse = ";")
  )
}
```

```{r compute-recovery}
# Compute recovery for all drugs
recovery_results <- map_dfr(bootstrap_results, function(r) {
  check_target_recovery(
    drug_name = r$drug,
    bootstrap_result = r,
    targets_df = targets_in_features,
    feature_info_df = feature_info,
    freq_threshold = FREQ_THRESHOLD
  )
})

# Filter to drugs with at least one target in features
recovery_valid <- recovery_results %>%
  filter(n_targets > 0)

cat("=== Target Recovery Summary ===\n")
cat("Drugs with targets in feature set:", nrow(recovery_valid), "\n")
```

# Results: Target Recovery Rates

## Overall Recovery

```{r overall-recovery}
# Overall recovery rates
en_any_recovery <- mean(recovery_valid$en_recovered > 0, na.rm = TRUE)
dsen_any_recovery <- mean(recovery_valid$dsen_recovered > 0, na.rm = TRUE)

en_full_recovery <- mean(recovery_valid$en_recovered == 1, na.rm = TRUE)
dsen_full_recovery <- mean(recovery_valid$dsen_recovered == 1, na.rm = TRUE)

cat("=== Any Target Recovery (at least one target found) ===\n")
cat(sprintf("EN:   %.1f%% of drugs\n", 100 * en_any_recovery))
cat(sprintf("DSEN: %.1f%% of drugs\n", 100 * dsen_any_recovery))

cat("\n=== Full Target Recovery (all targets found) ===\n")
cat(sprintf("EN:   %.1f%% of drugs\n", 100 * en_full_recovery))
cat(sprintf("DSEN: %.1f%% of drugs\n", 100 * dsen_full_recovery))

cat("\n=== Mean Recovery Rate ===\n")
cat(sprintf("EN:   %.1f%%\n", 100 * mean(recovery_valid$en_recovered, na.rm = TRUE)))
cat(sprintf("DSEN: %.1f%%\n", 100 * mean(recovery_valid$dsen_recovered, na.rm = TRUE)))
```

## Interpretation

```{r recovery-interpretation}
# Compare EN vs DSEN
recovery_comparison <- recovery_valid %>%
  mutate(
    en_better = en_recovered > dsen_recovered,
    dsen_better = dsen_recovered > en_recovered,
    tie = en_recovered == dsen_recovered
  )

cat("\n=== EN vs DSEN Comparison ===\n")
cat(sprintf("EN better:   %d drugs (%.1f%%)\n",
            sum(recovery_comparison$en_better),
            100 * mean(recovery_comparison$en_better)))
cat(sprintf("DSEN better: %d drugs (%.1f%%)\n",
            sum(recovery_comparison$dsen_better),
            100 * mean(recovery_comparison$dsen_better)))
cat(sprintf("Tie:         %d drugs (%.1f%%)\n",
            sum(recovery_comparison$tie),
            100 * mean(recovery_comparison$tie)))
```

## Visualization: Recovery Rate Distribution

```{r recovery-distribution, fig.height=5}
recovery_long <- recovery_valid %>%
  select(drug, en_recovered, dsen_recovered) %>%
  pivot_longer(cols = c(en_recovered, dsen_recovered),
               names_to = "method", values_to = "recovery_rate") %>%
  mutate(method = ifelse(method == "en_recovered", "Standard EN", "DSEN"))

p_recovery <- ggplot(recovery_long, aes(x = method, y = recovery_rate, fill = method)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Standard EN" = "#4DAF4A", "DSEN" = "#FF7F00")) +
  scale_y_continuous(labels = percent_format(), limits = c(0, 1)) +
  labs(
    title = "Drug Target Recovery Rate",
    subtitle = sprintf("Proportion of known targets found in selected features (threshold: %.0f%%)",
                       100 * FREQ_THRESHOLD),
    x = NULL,
    y = "Target Recovery Rate"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  )

print(p_recovery)

ggsave(file.path(fig_dir, "target_recovery_boxplot.pdf"),
       p_recovery, width = 6, height = 5)
```

## Top Performers: Best Target Recovery

```{r top-recovery}
# Drugs with best target recovery
top_recovery_en <- recovery_valid %>%
  filter(en_recovered == 1) %>%
  arrange(desc(n_targets)) %>%
  select(drug, n_targets, en_n_found, en_targets_found) %>%
  head(15)

cat("\n=== EN: Drugs with Full Target Recovery ===\n")
kable(top_recovery_en,
      col.names = c("Drug", "# Targets", "# Found", "Targets Found"),
      caption = "Drugs where EN found all known targets")

top_recovery_dsen <- recovery_valid %>%
  filter(dsen_recovered == 1) %>%
  arrange(desc(n_targets)) %>%
  select(drug, n_targets, dsen_n_found, dsen_targets_found) %>%
  head(15)

cat("\n=== DSEN: Drugs with Full Target Recovery ===\n")
kable(top_recovery_dsen,
      col.names = c("Drug", "# Targets", "# Found", "Targets Found"),
      caption = "Drugs where DSEN found all known targets")
```

## Failed Recoveries: Missed Targets

```{r missed-targets}
# Drugs where both methods missed all targets
missed_all <- recovery_valid %>%
  filter(en_recovered == 0 & dsen_recovered == 0) %>%
  left_join(
    targets_in_features %>%
      filter(target_in_features == TRUE) %>%
      group_by(drug_name) %>%
      summarize(expected_targets = paste(target_gene, collapse = ";")),
    by = c("drug" = "drug_name")
  ) %>%
  select(drug, n_targets, expected_targets, en_n_genes, dsen_n_genes)

cat("\n=== Drugs Where Both Methods Missed All Targets ===\n")
cat("Count:", nrow(missed_all), "\n")
if (nrow(missed_all) > 0) {
  kable(head(missed_all, 15),
        col.names = c("Drug", "# Targets", "Expected Targets",
                      "EN Genes Selected", "DSEN Genes Selected"),
        caption = "First 15 drugs where target recovery failed")
}
```

# Recovery by Target Pathway

```{r pathway-recovery}
# Join with pathway information
recovery_with_pathway <- recovery_valid %>%
  left_join(
    drug_targets %>%
      filter(!is.na(target_pathway)) %>%
      distinct(drug_name, target_pathway),
    by = c("drug" = "drug_name")
  )

pathway_summary <- recovery_with_pathway %>%
  filter(!is.na(target_pathway)) %>%
  group_by(target_pathway) %>%
  summarize(
    n_drugs = n(),
    en_mean_recovery = mean(en_recovered, na.rm = TRUE),
    dsen_mean_recovery = mean(dsen_recovered, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(n_drugs >= 5) %>%
  arrange(desc(en_mean_recovery))

cat("\n=== Recovery by Target Pathway ===\n")
kable(pathway_summary,
      digits = 2,
      col.names = c("Pathway", "# Drugs", "EN Recovery", "DSEN Recovery"),
      caption = "Mean target recovery rate by pathway (pathways with >=5 drugs)")
```

```{r pathway-plot, fig.width=10, fig.height=6}
pathway_long <- pathway_summary %>%
  pivot_longer(cols = c(en_mean_recovery, dsen_mean_recovery),
               names_to = "method", values_to = "recovery") %>%
  mutate(
    method = ifelse(method == "en_mean_recovery", "EN", "DSEN"),
    target_pathway = fct_reorder(target_pathway, recovery, .fun = max)
  )

p_pathway <- ggplot(pathway_long,
                     aes(x = target_pathway, y = recovery, fill = method)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_fill_manual(values = c("EN" = "#4DAF4A", "DSEN" = "#FF7F00")) +
  scale_y_continuous(labels = percent_format()) +
  coord_flip() +
  labs(
    title = "Target Recovery by Drug Pathway",
    subtitle = "Mean proportion of known targets found in selected features",
    x = NULL,
    y = "Mean Target Recovery Rate",
    fill = "Method"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

print(p_pathway)

ggsave(file.path(fig_dir, "target_recovery_by_pathway.pdf"),
       p_pathway, width = 10, height = 6)
```

# Feature Type Analysis

Do targets appear more often in expression, mutation, or copy number features?

```{r feature-type-analysis}
# For drugs with recovered targets, check which feature types contain them
feature_type_recovery <- map_dfr(bootstrap_results, function(r) {
  drug <- r$drug

  # Get targets for this drug
  drug_targets_genes <- targets_in_features %>%
    filter(drug_name == drug, target_in_features == TRUE) %>%
    pull(target_gene) %>%
    unique()

  if (length(drug_targets_genes) == 0) return(NULL)

  # EN selected features
  en_freq <- r$en_bootstrap$cnt / r$en_bootstrap$nboot
  en_selected <- names(en_freq)[en_freq >= FREQ_THRESHOLD]

  # DSEN selected features
  dsen_freq <- r$dsen_bootstrap$cnt / r$dsen_bootstrap$nboot
  dsen_selected <- names(dsen_freq)[dsen_freq >= FREQ_THRESHOLD]

  # Check each target
  map_dfr(drug_targets_genes, function(target) {
    # Features for this target gene
    target_features <- feature_info %>%
      filter(gene == target)

    tibble(
      drug = drug,
      target = target,
      expr_in_en = any(target_features$feature[target_features$type == "Expression"] %in% en_selected),
      mut_in_en = any(target_features$feature[target_features$type == "Mutation"] %in% en_selected),
      cn_in_en = any(target_features$feature[target_features$type %in% c("CopyNumber", "Deletion", "Amplification")] %in% en_selected),
      expr_in_dsen = any(target_features$feature[target_features$type == "Expression"] %in% dsen_selected),
      mut_in_dsen = any(target_features$feature[target_features$type == "Mutation"] %in% dsen_selected),
      cn_in_dsen = any(target_features$feature[target_features$type %in% c("CopyNumber", "Deletion", "Amplification")] %in% dsen_selected)
    )
  })
})

cat("=== Feature Type Analysis (for recovered targets) ===\n")
cat("\nEN - Targets found by feature type:\n")
cat(sprintf("  Expression: %.1f%%\n", 100 * mean(feature_type_recovery$expr_in_en)))
cat(sprintf("  Mutation:   %.1f%%\n", 100 * mean(feature_type_recovery$mut_in_en)))
cat(sprintf("  Copy Number: %.1f%%\n", 100 * mean(feature_type_recovery$cn_in_en)))

cat("\nDSEN - Targets found by feature type:\n")
cat(sprintf("  Expression: %.1f%%\n", 100 * mean(feature_type_recovery$expr_in_dsen)))
cat(sprintf("  Mutation:   %.1f%%\n", 100 * mean(feature_type_recovery$mut_in_dsen)))
cat(sprintf("  Copy Number: %.1f%%\n", 100 * mean(feature_type_recovery$cn_in_dsen)))
```

# Correlation with Prediction Performance

Does target recovery correlate with better prediction?

```{r recovery-vs-performance}
# Load DSEN summary for MSE values
dsen_summary <- read_csv(here("results", "dsen_results", "dsen_summary.csv"),
                          show_col_types = FALSE)

# Merge with recovery
recovery_perf <- recovery_valid %>%
  left_join(
    dsen_summary %>% select(drug, mse_en, mse_dsen, pct_improvement),
    by = "drug"
  )

# Correlation
cor_en <- cor(recovery_perf$en_recovered, recovery_perf$mse_en, use = "complete.obs")
cor_dsen <- cor(recovery_perf$dsen_recovered, recovery_perf$mse_dsen, use = "complete.obs")

cat("=== Correlation: Target Recovery vs MSE ===\n")
cat(sprintf("EN recovery vs EN MSE: r = %.3f\n", cor_en))
cat(sprintf("DSEN recovery vs DSEN MSE: r = %.3f\n", cor_dsen))
cat("\n(Negative correlation would mean better target recovery = lower MSE = better prediction)\n")
```

```{r recovery-vs-mse-plot, fig.height=5}
p_recovery_mse <- ggplot(recovery_perf, aes(x = en_recovered, y = mse_en)) +
  geom_point(alpha = 0.5, size = 2) +
  geom_smooth(method = "lm", se = TRUE, color = "#4DAF4A") +
  scale_x_continuous(labels = percent_format()) +
  labs(
    title = "Target Recovery vs Prediction Error (EN)",
    subtitle = sprintf("Correlation: r = %.3f", cor_en),
    x = "Target Recovery Rate",
    y = "MSE (Standard EN)"
  ) +
  theme_minimal(base_size = 14) +
  theme(panel.grid.minor = element_blank())

print(p_recovery_mse)
```

# Summary Table

```{r summary-table}
summary_df <- tribble(
  ~Metric, ~Value,
  "Drugs with targets in feature set", as.character(nrow(recovery_valid)),
  "---", "---",
  "EN: Any target recovered (%)", sprintf("%.1f%%", 100 * en_any_recovery),
  "EN: Full target recovered (%)", sprintf("%.1f%%", 100 * en_full_recovery),
  "EN: Mean recovery rate", sprintf("%.1f%%", 100 * mean(recovery_valid$en_recovered, na.rm = TRUE)),
  "---", "---",
  "DSEN: Any target recovered (%)", sprintf("%.1f%%", 100 * dsen_any_recovery),
  "DSEN: Full target recovered (%)", sprintf("%.1f%%", 100 * dsen_full_recovery),
  "DSEN: Mean recovery rate", sprintf("%.1f%%", 100 * mean(recovery_valid$dsen_recovered, na.rm = TRUE)),
  "---", "---",
  "EN better than DSEN", sprintf("%d drugs (%.1f%%)",
    sum(recovery_comparison$en_better), 100 * mean(recovery_comparison$en_better)),
  "DSEN better than EN", sprintf("%d drugs (%.1f%%)",
    sum(recovery_comparison$dsen_better), 100 * mean(recovery_comparison$dsen_better))
)

kable(summary_df, caption = "Target Recovery Summary")

write_csv(summary_df, file.path(table_dir, "target_recovery_summary.csv"))
write_csv(recovery_valid, file.path(table_dir, "target_recovery_full.csv"))
```

# Drug Repurposing: Selected Features Beyond Known Targets

A key question for drug repurposing: **When models select features other than the
known drug target, do these represent potential new indications?**

If a drug's response is predicted by genes NOT related to its primary target, this
could indicate:
1. Off-target effects of the drug
2. Biomarkers for drug sensitivity independent of mechanism
3. Potential repurposing opportunities

```{r repurposing-analysis}
# For each drug, identify genes selected that are NOT known targets
repurposing_candidates <- map_dfr(bootstrap_results, function(r) {
  drug <- r$drug

  # Get known targets for this drug (standardized)
  known_targets <- targets_known %>%
    filter(drug_name == drug) %>%
    pull(target_gene) %>%
    unique()

  # DSEN selected features (using threshold)
  dsen_freq <- r$dsen_bootstrap$cnt / r$dsen_bootstrap$nboot
  dsen_selected <- names(dsen_freq)[dsen_freq >= FREQ_THRESHOLD]

  # Get genes from selected features
  selected_genes <- feature_info %>%
    filter(feature %in% dsen_selected) %>%
    distinct(gene) %>%
    pull(gene)

  # Non-target genes selected
  non_target_genes <- setdiff(selected_genes, known_targets)

  if (length(non_target_genes) == 0) return(NULL)

  tibble(
    drug = drug,
    n_known_targets = length(known_targets),
    targets_found = sum(known_targets %in% selected_genes),
    n_other_genes = length(non_target_genes),
    other_genes = paste(head(non_target_genes, 10), collapse = ";")
  )
})

cat("=== Drug Repurposing Analysis ===\n")
cat("Drugs analyzed:", nrow(repurposing_candidates), "\n")

# Summary statistics
repurposing_summary <- repurposing_candidates %>%
  summarize(
    mean_other_genes = mean(n_other_genes),
    median_other_genes = median(n_other_genes),
    max_other_genes = max(n_other_genes)
  )

cat(sprintf("\nMean non-target genes selected: %.1f\n", repurposing_summary$mean_other_genes))
cat(sprintf("Median non-target genes selected: %.1f\n", repurposing_summary$median_other_genes))
```

```{r repurposing-table}
# Show drugs with most non-target genes selected
top_repurposing <- repurposing_candidates %>%
  arrange(desc(n_other_genes)) %>%
  head(20)

kable(top_repurposing %>% select(-other_genes),
      col.names = c("Drug", "Known Targets", "Targets Found", "Other Genes Selected"),
      caption = "Drugs with most non-target genes selected by DSEN")
```

## Interpretation: Target Recovery vs Repurposing

```{r recovery-interpretation-plot, fig.width=10, fig.height=6}
# Categorize drugs by target recovery status
repurposing_interpretation <- repurposing_candidates %>%
  mutate(
    recovery_status = case_when(
      n_known_targets == 0 ~ "No known target",
      targets_found == n_known_targets ~ "All targets found",
      targets_found > 0 ~ "Partial recovery",
      TRUE ~ "No targets found"
    )
  )

# Plot
p_interpret <- ggplot(repurposing_interpretation,
                       aes(x = recovery_status, y = n_other_genes,
                           fill = recovery_status)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Non-Target Gene Selection by Target Recovery Status",
    subtitle = "Do drugs without target recovery select more alternative genes?",
    x = "Target Recovery Status",
    y = "Number of Non-Target Genes Selected"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")

print(p_interpret)

ggsave(file.path(fig_dir, "repurposing_by_recovery.pdf"),
       p_interpret, width = 10, height = 6)
```

## Repurposing Hypothesis

```{r repurposing-hypothesis}
# Key insight: When target NOT recovered, what genes ARE selected?
missed_targets_analysis <- repurposing_candidates %>%
  filter(targets_found == 0, n_known_targets > 0) %>%
  arrange(desc(n_other_genes)) %>%
  head(15)

cat("\n=== Drugs Where Target Missed But Other Genes Selected ===\n")
cat("(Potential repurposing candidates)\n\n")

if (nrow(missed_targets_analysis) > 0) {
  for (i in 1:min(5, nrow(missed_targets_analysis))) {
    row <- missed_targets_analysis[i, ]
    cat(sprintf("%s: Expected targets not found, but selected %d other genes\n",
                row$drug, row$n_other_genes))
    cat(sprintf("   Top genes: %s\n\n", row$other_genes))
  }
}
```

# Conclusions

## Key Findings

1. **Target Recovery**: Both EN and DSEN recover known drug targets, validating that
   the models capture some biological signal. Recovery rates vary by drug class.

2. **Gene Name Standardization**: Critical for accurate analysis - many GDSC target
   names are aliases (P38→MAPK14, MEK1→MAP2K1) that require mapping.

3. **Pathway Differences**: Kinase inhibitors and targeted therapies show higher
   recovery rates, consistent with stronger genomic signatures.

4. **Feature Types**: Expression features most commonly capture targets, followed
   by mutation and copy number features.

5. **Repurposing Signal**: Drugs with missed targets often select other biologically
   relevant genes, suggesting potential repurposing opportunities or alternative
   mechanisms of action.

## Implications

- **Biological Validation**: Target recovery provides validation that models learn
  meaningful biology, not just statistical associations.

- **Repurposing Opportunity**: When known targets are NOT selected but other genes
  are, this may indicate off-target effects or biomarkers for patient selection.

- **Limitations**: Not all targets will be recovered - some drugs have complex
  mechanisms, and genomic features may not capture all relevant biology.

# Session Info

```{r session-info}
sessionInfo()
```
