---
title: "Tissue-Specific Target Recovery: DSEN Shared vs Tissue-Specific Blocks"
author: "Amir Asiaee"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6,
  dpi = 150
)
```

# Overview

This analysis extends the target recovery analysis from `10_drug_target_validation.Rmd`
by incorporating **tissue-specific** DSEN coefficients. The key question: does DSEN's
tissue-specific block capture drug targets that the shared block misses?

Key improvements over experiment 10:

1. **Corrected gene extraction**: handles all mutation suffixes (`_mut.hs`, `_mut.tr`,
   `_mut.mis`, `_mut.hsrec`) plus bare RPPA protein features
2. **Tissue-specific recovery**: uses proper bootstrap frequency counts from
   `tissue_cnt` (not averaged coefficients)
3. **Combined recovery**: union of shared + tissue-specific gene sets

# Setup

```{r load-packages}
library(here)
library(tidyverse)
library(knitr)
library(patchwork)
library(scales)
```

```{r load-paths}
source(here("methods", "00_paths.R"))
fig_dir <- file.path(paths$results, "figures")
table_dir <- file.path(paths$results, "tables")
```

# Feature-to-Gene Mapping

## Feature Type Inventory

The predictor matrix has 86,546 features. We must correctly map each to a gene symbol.

```{r load-feature-names}
load(file.path(paths$clean, "ccle.Rda"))
feature_names <- colnames(x)
cat("Total features:", length(feature_names), "\n")
rm(x, y, cls)
```

```{r feature-type-inventory}
# Classify features by type
feature_info <- tibble(feature = feature_names) %>%
  mutate(
    type = case_when(
      grepl("_CN$", feature) ~ "CopyNumber",
      grepl("_mut\\.mis$", feature) ~ "Mutation_missense",
      grepl("_Expr$", feature) & !grepl("_\\d+_at_Expr$", feature) ~ "Expression",
      grepl("_mut\\.tr$", feature) ~ "Mutation_truncating",
      grepl("_mut\\.hs$", feature) ~ "Mutation_highsig",
      grepl("_mut\\.hsrec$", feature) ~ "Mutation_recurrent",
      grepl("_\\d+_at_Expr$", feature) ~ "Probe_expression",
      grepl("_Caution$", feature) ~ "QualityFlag",
      grepl("_ValidationFailed$|_ValidationUnavailable$", feature) ~ "QualityFlag",
      TRUE ~ "RPPA"
    )
  )

cat("Feature type counts:\n")
print(table(feature_info$type))
```

## Gene Extraction from Standard Features

```{r gene-extraction}
# Extract gene names using correct regex for ALL mutation suffixes
feature_info <- feature_info %>%
  mutate(
    gene = case_when(
      # Standard genomic features: remove known suffixes
      type %in% c("CopyNumber", "Expression",
                   "Mutation_missense", "Mutation_truncating",
                   "Mutation_highsig", "Mutation_recurrent") ~
        sub("_mut\\.hsrec$|_mut\\.hs$|_mut\\.tr$|_mut\\.mis$|_Expr$|_CN$", "", feature),
      # Probe-level and quality flags: no gene mapping
      type %in% c("Probe_expression", "QualityFlag") ~ NA_character_,
      # RPPA: handled separately below
      type == "RPPA" ~ NA_character_
    )
  )

cat("Features with gene mapping (before RPPA):",
    sum(!is.na(feature_info$gene)), "\n")
cat("Features without mapping (probes + quality flags):",
    sum(is.na(feature_info$gene) & feature_info$type != "RPPA"), "\n")
cat("RPPA features to map:", sum(feature_info$type == "RPPA"), "\n")
```

## RPPA Protein-to-Gene Mapping

139 RPPA features use protein/antibody names. We map each to its HGNC gene symbol.

```{r load-rppa-mapping}
rppa_mapping <- read_csv(here("data", "rppa_gene_mapping.csv"), show_col_types = FALSE)
cat("RPPA mapping loaded:", nrow(rppa_mapping), "entries\n")
cat("Unique genes:", n_distinct(rppa_mapping$hgnc_symbol), "\n")

# Show mapping examples
kable(
  rppa_mapping %>%
    filter(rppa_feature != hgnc_symbol) %>%
    select(rppa_feature, hgnc_symbol, notes) %>%
    head(20),
  caption = "RPPA protein-to-gene mapping examples"
)
```

```{r apply-rppa-mapping}
# Apply RPPA mapping
rppa_lookup <- setNames(rppa_mapping$hgnc_symbol, rppa_mapping$rppa_feature)

feature_info <- feature_info %>%
  mutate(
    gene = ifelse(type == "RPPA" & feature %in% names(rppa_lookup),
                  rppa_lookup[feature],
                  gene)
  )

# Verify: how many RPPA features mapped?
rppa_mapped <- sum(!is.na(feature_info$gene[feature_info$type == "RPPA"]))
cat("RPPA features mapped:", rppa_mapped, "/",
    sum(feature_info$type == "RPPA"), "\n")

# Overall mapping summary
cat("\nOverall mapping summary:\n")
feature_info %>%
  group_by(type) %>%
  summarize(
    n = n(),
    mapped = sum(!is.na(gene)),
    pct = round(100 * mapped / n, 1)
  ) %>%
  kable()
```

```{r gene-summary}
# Unique genes available for target matching
genes_in_data <- unique(feature_info$gene[!is.na(feature_info$gene)])
cat("Unique genes in feature set:", length(genes_in_data), "\n")
```

# Load Drug Targets

```{r load-targets}
drug_targets <- read_csv(here("data", "drug_targets_curated.csv"), show_col_types = FALSE)
cat("Drug target entries:", nrow(drug_targets), "\n")
cat("Unique drugs:", n_distinct(drug_targets$drug_name), "\n")
```

## Gene Name Standardization

```{r standardize-genes}
gene_mapping <- read_csv(here("data", "gene_name_mapping.csv"), show_col_types = FALSE)

standardize_gene <- function(gene, mapping_df) {
  match_idx <- match(toupper(gene), toupper(mapping_df$alias))
  if (!is.na(match_idx)) {
    return(mapping_df$hgnc_symbol[match_idx])
  }
  return(gene)
}

targets_known <- drug_targets %>%
  filter(!is.na(target_gene), source != "not_found") %>%
  mutate(
    original_target = target_gene,
    target_gene = sapply(target_gene, standardize_gene, mapping_df = gene_mapping),
    target_in_features = target_gene %in% genes_in_data
  )

cat("Drugs with gene targets:", n_distinct(targets_known$drug_name), "\n")
cat("Drugs with at least one target in feature set:",
    n_distinct(targets_known$drug_name[targets_known$target_in_features]), "\n")
```

# Load Bootstrap Results

```{r load-bootstrap}
# Load v2 bootstrap results (threshold_25_v2 with bug fixes)
bootstrap_file <- here("results", "dsen_results", "threshold_25_v2", "dsen_bootstrap_results.Rda")
stopifnot(file.exists(bootstrap_file))
load(bootstrap_file)
bootstrap_results <- all_results
rm(all_results)
cat("Loaded bootstrap results for", length(bootstrap_results), "drugs\n")

# Verify tissue_cnt exists
r1 <- bootstrap_results[[1]]
has_tissue_cnt <- "tissue_cnt" %in% names(r1$dsen_bootstrap)
cat("Has tissue_cnt:", has_tissue_cnt, "\n")
if (has_tissue_cnt) {
  cat("tissue_cnt dimensions:", dim(r1$dsen_bootstrap$tissue_cnt), "\n")
}
cat("nboot:", r1$dsen_bootstrap$nboot, "\n")
```

# Target Recovery Analysis

## Helper Function

```{r map-features-to-genes}
# Map a vector of feature names to gene symbols
map_features_to_genes <- function(features, feature_info_df) {
  feature_info_df %>%
    filter(feature %in% features, !is.na(gene)) %>%
    pull(gene) %>%
    unique()
}
```

## Compute Recovery at Multiple Thresholds

```{r compute-recovery}
THRESHOLDS <- c(0.25, 0.50, 0.75)

compute_all_recovery <- function(bootstrap_results, targets_known_df,
                                  feature_info_df, thresholds) {

  results_list <- list()

  for (thresh in thresholds) {
    cat(sprintf("\n=== Threshold: %.2f ===\n", thresh))

    drug_results <- map_dfr(bootstrap_results, function(r) {
      drug <- r$drug
      if (r$status != "success") return(NULL)

      # Get targets for this drug
      drug_tgts <- targets_known_df %>%
        filter(drug_name == drug, target_in_features == TRUE) %>%
        pull(target_gene) %>%
        unique()

      if (length(drug_tgts) == 0) {
        return(tibble(
          drug = drug,
          threshold = thresh,
          n_targets = 0,
          en_any = NA, en_full = NA,
          dsen_shared_any = NA, dsen_shared_full = NA,
          dsen_tissue_any = NA, dsen_tissue_full = NA,
          dsen_combined_any = NA, dsen_combined_full = NA,
          en_targets_found = NA_character_,
          dsen_shared_targets = NA_character_,
          dsen_tissue_targets = NA_character_,
          dsen_combined_targets = NA_character_
        ))
      }

      # --- EN ---
      en_freq <- r$en_bootstrap$cnt / r$en_bootstrap$nboot
      en_selected <- names(en_freq)[en_freq >= thresh]
      en_genes <- map_features_to_genes(en_selected, feature_info_df)
      en_found <- intersect(drug_tgts, en_genes)

      # --- DSEN shared ---
      dsen_freq <- r$dsen_bootstrap$cnt / r$dsen_bootstrap$nboot
      dsen_shared_selected <- names(dsen_freq)[dsen_freq >= thresh]
      dsen_shared_genes <- map_features_to_genes(dsen_shared_selected, feature_info_df)
      dsen_shared_found <- intersect(drug_tgts, dsen_shared_genes)

      # --- DSEN tissue-specific ---
      dsen_tissue_genes <- character(0)
      dsen_tissue_selected <- character(0)
      if ("tissue_cnt" %in% names(r$dsen_bootstrap)) {
        tissue_cnt <- r$dsen_bootstrap$tissue_cnt
        nboot <- r$dsen_bootstrap$nboot
        tissue_freq <- tissue_cnt / nboot
        # Feature selected if freq >= threshold in ANY tissue
        tissue_any <- rownames(tissue_freq)[
          apply(tissue_freq >= thresh, 1, any)
        ]
        dsen_tissue_selected <- tissue_any
        dsen_tissue_genes <- map_features_to_genes(tissue_any, feature_info_df)
      }
      dsen_tissue_found <- intersect(drug_tgts, dsen_tissue_genes)

      # --- DSEN combined ---
      dsen_combined_genes <- union(dsen_shared_genes, dsen_tissue_genes)
      dsen_combined_found <- intersect(drug_tgts, dsen_combined_genes)

      tibble(
        drug = drug,
        threshold = thresh,
        n_targets = length(drug_tgts),
        en_n_found = length(en_found),
        dsen_shared_n_found = length(dsen_shared_found),
        dsen_tissue_n_found = length(dsen_tissue_found),
        dsen_combined_n_found = length(dsen_combined_found),
        en_any = length(en_found) > 0,
        en_full = length(en_found) == length(drug_tgts),
        dsen_shared_any = length(dsen_shared_found) > 0,
        dsen_shared_full = length(dsen_shared_found) == length(drug_tgts),
        dsen_tissue_any = length(dsen_tissue_found) > 0,
        dsen_tissue_full = length(dsen_tissue_found) == length(drug_tgts),
        dsen_combined_any = length(dsen_combined_found) > 0,
        dsen_combined_full = length(dsen_combined_found) == length(drug_tgts),
        en_n_selected = length(en_selected),
        dsen_shared_n_selected = length(dsen_shared_selected),
        dsen_tissue_n_selected = length(dsen_tissue_selected),
        en_targets_found = paste(en_found, collapse = ";"),
        dsen_shared_targets = paste(dsen_shared_found, collapse = ";"),
        dsen_tissue_targets = paste(dsen_tissue_found, collapse = ";"),
        dsen_combined_targets = paste(dsen_combined_found, collapse = ";")
      )
    })

    results_list[[as.character(thresh)]] <- drug_results
  }

  bind_rows(results_list)
}

recovery_df <- compute_all_recovery(
  bootstrap_results, targets_known, feature_info, THRESHOLDS
)

# Filter to drugs with at least one target in features
recovery_valid <- recovery_df %>%
  filter(n_targets > 0)

cat("\nDrugs with targets in feature set:",
    n_distinct(recovery_valid$drug), "\n")
```

## Summary Statistics

```{r recovery-summary}
summary_df <- recovery_valid %>%
  group_by(threshold) %>%
  summarize(
    n_drugs = n_distinct(drug),
    en_any_pct = round(100 * mean(en_any, na.rm = TRUE), 1),
    en_any_n = sum(en_any, na.rm = TRUE),
    dsen_shared_any_pct = round(100 * mean(dsen_shared_any, na.rm = TRUE), 1),
    dsen_shared_any_n = sum(dsen_shared_any, na.rm = TRUE),
    dsen_tissue_any_pct = round(100 * mean(dsen_tissue_any, na.rm = TRUE), 1),
    dsen_tissue_any_n = sum(dsen_tissue_any, na.rm = TRUE),
    dsen_combined_any_pct = round(100 * mean(dsen_combined_any, na.rm = TRUE), 1),
    dsen_combined_any_n = sum(dsen_combined_any, na.rm = TRUE),
    .groups = "drop"
  )

cat("\n=== Target Recovery Summary (Any Target Found) ===\n\n")
kable(
  summary_df %>%
    select(threshold, n_drugs,
           en_any_pct, en_any_n,
           dsen_shared_any_pct, dsen_shared_any_n,
           dsen_tissue_any_pct, dsen_tissue_any_n,
           dsen_combined_any_pct, dsen_combined_any_n),
  col.names = c("Threshold", "Drugs", "EN %", "EN n",
                "Shared %", "Shared n",
                "Tissue %", "Tissue n",
                "Combined %", "Combined n"),
  caption = "Target recovery rates: any known target found among selected features"
)
```

## Tissue-Only Recoveries

Drugs where the target was found exclusively in tissue-specific blocks.

```{r tissue-only}
recovery_050 <- recovery_valid %>%
  filter(threshold == 0.50)

tissue_only <- recovery_050 %>%
  filter(!dsen_shared_any & dsen_tissue_any)

cat("=== Drugs with target found ONLY in tissue-specific block (threshold 0.50) ===\n")
cat("Count:", nrow(tissue_only), "\n\n")

if (nrow(tissue_only) > 0) {
  kable(
    tissue_only %>%
      select(drug, n_targets, dsen_tissue_targets),
    col.names = c("Drug", "# Targets", "Tissue-Specific Targets Found"),
    caption = "Targets recovered exclusively from tissue-specific coefficients"
  )
}
```

## Shared-Only Recoveries

Drugs where the target was found only in the shared block.

```{r shared-only}
shared_only <- recovery_050 %>%
  filter(dsen_shared_any & !dsen_tissue_any)

cat("=== Drugs with target found ONLY in shared block (threshold 0.50) ===\n")
cat("Count:", nrow(shared_only), "\n\n")

if (nrow(shared_only) > 0) {
  kable(
    shared_only %>%
      select(drug, n_targets, dsen_shared_targets),
    col.names = c("Drug", "# Targets", "Shared Targets Found"),
    caption = "Targets recovered exclusively from shared coefficients"
  )
}
```

## Both Blocks

```{r both-blocks}
both_blocks <- recovery_050 %>%
  filter(dsen_shared_any & dsen_tissue_any)

cat("=== Drugs with target found in BOTH blocks (threshold 0.50) ===\n")
cat("Count:", nrow(both_blocks), "\n\n")

if (nrow(both_blocks) > 0) {
  kable(
    both_blocks %>%
      select(drug, n_targets, dsen_shared_targets, dsen_tissue_targets),
    col.names = c("Drug", "# Targets", "Shared Targets", "Tissue Targets"),
    caption = "Targets recovered from both shared and tissue-specific coefficients"
  )
}
```

# Case Study Table

Key drugs illustrating the shared vs tissue-specific pattern.

```{r case-studies}
# Define key drugs of interest
key_drugs <- c("PLX-4720", "Dabrafenib", "Nutlin-3a(-)", "Nutlin-3a",
               "NVP-TAE684", "Quizartinib", "Sunitinib",
               "Navitoclax", "Ruxolitinib", "Fedratinib",
               "Afatinib", "Cetuximab")

case_studies <- recovery_050 %>%
  filter(drug %in% key_drugs | dsen_combined_any) %>%
  left_join(
    targets_known %>%
      filter(target_in_features) %>%
      group_by(drug_name) %>%
      summarize(known_targets = paste(unique(target_gene), collapse = "; "),
                .groups = "drop"),
    by = c("drug" = "drug_name")
  ) %>%
  mutate(
    en_status = ifelse(en_any, "Yes", "No"),
    shared_status = ifelse(dsen_shared_any, "Yes", "No"),
    tissue_status = ifelse(dsen_tissue_any, "Yes", "No"),
    combined_status = ifelse(dsen_combined_any, "Yes", "No")
  ) %>%
  select(drug, known_targets, en_status, shared_status, tissue_status,
         en_targets_found, dsen_shared_targets, dsen_tissue_targets) %>%
  arrange(desc(tissue_status), desc(shared_status))

cat("=== Case Studies: Target Recovery by Block ===\n")
kable(
  case_studies,
  col.names = c("Drug", "Known Targets", "EN", "Shared", "Tissue",
                "EN Found", "Shared Found", "Tissue Found"),
  caption = "Target recovery decomposition for key drugs (threshold = 0.50)"
)
```

# Comparison with Ad-Hoc Analysis

```{r adhoc-comparison}
cat("=== Proper Analysis vs Ad-Hoc (threshold 0.50) ===\n\n")

s <- summary_df %>% filter(threshold == 0.50)

cat(sprintf("EN:             %d/%d drugs (%.1f%%)\n",
            s$en_any_n, s$n_drugs, s$en_any_pct))
cat(sprintf("DSEN shared:    %d/%d drugs (%.1f%%)\n",
            s$dsen_shared_any_n, s$n_drugs, s$dsen_shared_any_pct))
cat(sprintf("DSEN tissue:    %d/%d drugs (%.1f%%)\n",
            s$dsen_tissue_any_n, s$n_drugs, s$dsen_tissue_any_pct))
cat(sprintf("DSEN combined:  %d/%d drugs (%.1f%%)\n",
            s$dsen_combined_any_n, s$n_drugs, s$dsen_combined_any_pct))

cat("\nAd-hoc estimates (for reference):\n")
cat("  EN: 4.9%, DSEN shared: 5.3%, DSEN tissue: 13.2%, Combined: 13.6%\n")
cat("\nNote: ad-hoc used non-zero mean coefficients; proper uses freq >= 0.50\n")
```

# Save Results

```{r save-results}
# Per-drug results (threshold 0.50 for main table)
v2_table_dir <- file.path(table_dir, "threshold_25_v2")
dir.create(v2_table_dir, showWarnings = FALSE, recursive = TRUE)
write_csv(recovery_050,
          file.path(v2_table_dir, "target_recovery_extended.csv"))

# Summary across thresholds
write_csv(summary_df,
          file.path(v2_table_dir, "target_recovery_extended_summary.csv"))

# Full R objects
v2_dsen_dir <- here("results", "dsen_results", "threshold_25_v2")
dir.create(v2_dsen_dir, showWarnings = FALSE, recursive = TRUE)
save(recovery_df, summary_df, feature_info, rppa_mapping,
     file = file.path(v2_dsen_dir, "tissue_target_recovery.Rda"))

cat("Results saved.\n")
```

# Session Info

```{r session-info}
sessionInfo()
```
